### 2021-08-11 manuscript/coverLetter/Letter.tex
updates to package str

\documentclass[12pt]{article}
\usepackage[letterpaper,margin=1cm,bottom=2cm,top=3cm,nohead]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{amsthm}
\newcommand{\email}[1]{\href{mailto:#1}{\tt \textcolor{cornflower}{#1}}}
\usepackage{xcolor}
\definecolor{plum}{rgb}{0.36078, 0.20784, 0.4}
\definecolor{chameleon}{rgb}{0.30588, 0.60392, 0.023529}
\definecolor{cornflower}{rgb}{0.12549, 0.29020, 0.52941}
\definecolor{scarlet}{rgb}{0.8, 0, 0}
\definecolor{brick}{rgb}{0.64314, 0, 0}
\usepackage[%
    colorlinks=true,  %display links in different colors
    linkcolor=cornflower, %internal links
    citecolor=scarlet, %references to a bibliography
    urlcolor=chameleon %URLs
]{hyperref}
\usepackage{graphicx}
\setlength{\parskip}{0.75em}
\setlength{\parindent}{1.75em}
\usepackage{afterpage}
\usepackage[strict]{changepage}
\pagestyle{empty}

\begin{document}

\begin{minipage}[c]{8in}\vskip-2cm
\begin{flushleft}
	\begin{minipage}[c]{4cm}
		\begin{flushleft}
			\includegraphics*[width=4cm]{lsu.png}%
		\end{flushleft}
	\end{minipage}
	\,\,
	\begin{minipage}[c]{0.25cm}
		\vrule depth 0.1\textheight
	\end{minipage}
	\,
	\begin{minipage}[c]{8cm}
		\footnotesize
		{\sffamily
			Tad Dallas \\
			Department of Biological Sciences \\
			Louisiana State University \\
			Baton Rouge, LA 70802 \\
			phone: +1 225 239 0410  \\
			\email{tad.a.dallas@gmail.com}
		}
	\end{minipage}
\end{flushleft}
\end{minipage}

\begin{adjustwidth}{4em}{4em}
~\vskip.4cm
\hfill \today
\vskip.5cm

{\noindent Dear \textit{Ecology Letters} editorial board,} \vskip4pt

Please find the enclosed manuscript ``A latitudinal signal in the relationship between species geographic range size and climatic niche area'', which we submit for consideration as a research article. \\

Species with wider climatic niche breadths are expected to be more widely distributed geographically. But while there appears to be strong evidence for this relationship, an understanding of when deviations would be expected as a result of geographic effects or species traits is an open question. Here, we use data on over 3000 species of mammal, bird, and tree, relating the model residuals from the relationship between geographic range size and climatic niche area to species traits and geographic position, and compare these residuals to what would be expected under a null model. \\

We find evidence for an effect of species latitudinal centre on model residuals, suggesting the existence of a consistent spatial signal in deviation from range size -- climatic niche area scaling relationships across a diverse set of species. Some species traits were also important (e.g., body size for mammals), but the distribution of model residuals comparing empirical to null relationship did not differ appreciably. Together, our analyses suggest that geographic range size -- climatic niche area relationships are more a function of the landscape and a spatially autocorrelated environment than driven by species traits. \\

We believe this article will be of interest to a broad range of ecologists and biogeographers, as it addresses a longstanding question in ecology concerning the link between species niche and distribution. In advance, we appreciate your consideration and assistance. Please don't hesitate to contact us with any questions.

\changepage{}{}{}{}{}{1cm}{}{}{}

\vskip1cm
\hfill
\noindent\begin{minipage}[b]{7cm}
    Sincerely,\\ ~
    Tad Dallas
\end{minipage}

\end{adjustwidth}
\end{document}

### 2021-08-11 manuscript/manuscript.tex
updates to package str

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{a4paper}
\usepackage[sort,comma]{natbib}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{setspace}
\usepackage{rotating}
\usepackage{pdflscape}
\usepackage[most]{tcolorbox}

\doublespacing

\title{\large \texttt{insectDisease}: programmatic access to the \textit{Ecological Database of the World's Insect Pathogens}}

\renewcommand\Authfont{\fontsize{12}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{10}{10.8}\itshape}

\author[a,b,*]{Tad A Dallas}
\author[c]{Colin Carlson}

% \author[c]{David Onstad}

\affil[a]{Department of Biological Sciences, Louisiana State University, Baton Rouge, LA, USA}
\affil[b]{Department of Biological Sciences, University of South Carolina, Columbia, SC, USA}

\affil[c]{  }
\affil[d]{  }

\renewcommand\Authands{ and }
 \date{ \small *Corresponding author: tad.a.dallas@gmail.com}

%%% Target journal
% Ecography (software note)
%
%

%%% Potential reviewers
% 
% 
% 
%

%%% Editors
% Thiago Rangel
% 
%

\begin{document}
\maketitle
\setstretch{1}

\vspace{-1cm}
\noindent \textbf{Author contributions}: TAD developed the R package, with taxonomic functions developed by CC. All authors contributed to manuscript writing.  \\

\noindent \textbf{Acknowledgements}: We sincerely thank all individuals associated with the collection and curation of the \textit{Ecological Database of the World's Insect Pathogens} data, specifically David Onstad at the US Department of Agriculture. \\

\noindent \textbf{Conflict of interest}: The authors have no conflicts of interest to declare. \\

\noindent \textbf{Keywords}: insect pathogens, crop pest, experimental infection,   \\

\clearpage
\setstretch{1.35}
\linenumbers

\subsection*{Abstract}

\medskip

\noindent \textbf{Running title}: Ecological Database of the World's Insect Pathogens (EDWIP)  \\

\clearpage
\setstretch{1.9}
\section*{Introduction}

% the lack of insect disease data
There are a number of data sources documenting host-pathogen associations, especially for pathogens of mammals \citep{gibb2021, patrick2017}, birds \citep{bensch2009}, and fish \citep{fishpest}. These have been incredibly important to our understanding of what determines pathogen host range, pathogen species richness across a set of hosts, and overall host-pathogen network structure (e.g, \citep{dallas2018, carlson2020}). But while some host groups are well-studied, there are taxonomic gaps in our understanding of host-pathogen associations. For instance, insect host-pathogen relationships have considerably less open-source data available, despite their inherent importance to agricultural crops and vector-borne disease, in addition to the numerical dominance of insect species over other taxa \citep{stork2015}. This is a clear knowledge gap.

% the file drawer issue and
\paragraph*{}
Additionally, all of the previous data sources listed before have dedicated researchers and government agencies responsible for data deposition and curation in openly accessible formats. However, some data have not been as lucky, at no fault of the original data curators. These data run the risk of disappearing into a file drawer or on an external hard drive, potentially shared with a small number of researchers but not accessible to the scientific community at large. One data resource arguably close to this point of disappearance is the \textit{Ecological Database of the World's Insect Pathogens} (EDWIP) \citep{onstad1997}.

% edwip
\paragraph*{}
The EDWIP data consist of experimental infections and field observations of the interactions between insect hosts and a number of nematode, viral, protozoan, fungal, and bacterial pathogens \citep{braxton2003}. In the EDWIP data, a number of pathogen infection types were excluded, including vector-borne pathogens which do not infect the insect host. An interesting component of EDWIP is the existence of negative associations -- attempts to inoculate a host with a given pathogen that failed to infect -- for some host groups (Figure \ref{fig:bubble}). Failed infections represent \textit{true} absences or incompatibilities between a given host and pathogen. These data are incredibly useful to pathogen host range and host-pathogen interaction modeling, but we rarely have data on these known non-interactions. Initially created in 1996, the data have been dynamically updated over time. However, it is unclear how long this curation continued, and how many different versions of the data may be in existence without proper versioning. To this point, the numbers of insect-host pathogen interactions differs slightly from previously published versions of the data \citep{braxton2003}, but these changes are relatively minor.

\subsection*{Solution statement}

To preserve these data in a format that is well-documented, openly accessible, versioned, and flexible for continued development, we created the \texttt{insectDisease} R package. In doing so, we implicitly adhere to the FAIR (Findable, Accessible, Interoperable, Resuable) guidelines for managing data. By hosting the data openly on GitHub, and versioning releases of the data with a permanent identifier (DOI), we ensure the longevity and versioned curation of this data resource. Finally, the incorporation of taxonomic data through \texttt{taxize} \citep{chamberlain2013} makes sure that host and pathogen taxonomic names are updated.

\subsection*{Data specification}

% Discuss more of the in-depth view of the R package as a means for data delivery, and what the overall workflow looks like. Discuss how it's open and all part of the Verena ecosystem (or whatever). Flex on data size and point to figures displaying the positive and negative associations and such.

\paragraph*{Package structure}

Data products are broken down by pathogen group; nematodes (\texttt{data(nematode)}), viruses (\texttt{data(viruses)}), and non-viral pathogens, which include protozoan, fungi, and bacteria (\texttt{data(nvpassoc)}). Data on negative associations is stored collectively instead of being delineated by pathogen group (\texttt{data(negative)}), but information on pathogen group is provided within each of these files, allowing for sorting of negative interactions based on the initial pathogen groupings.

\paragraph*{}
Each of the pathogen groups slightly differs in the available data on experimental infections. For instance, nematode infections contain information on soil type and associated bacteria, virus infection data has information on viral dose, and non-viral pathogens (protozoans, fungi, and bacteria) have information on intermediate host species.

\paragraph*{}
Data are also available on the insect host species themselves (e.g., \texttt{?hosts} or \texttt{data(hosts)}). These data contain some information on Canadian province where the host is found (\texttt{ProvinceI} column), what it eats (\texttt{Food} column), and what type of habitat it is found in (\texttt{Habitat} column). Additionally, a column on host insect pest status is present, offering the opportunity to explore study effort and pathogen specificity dependent on the pest status of the insect host.

\paragraph*{Metadata and package documentation}

Differences in features across the data on different pathogen types (e.g., \texttt{?nematodes} relative to \texttt{?viruses}) make combining these data difficult without a loss of information. However, there is clear utility in having all of the data in a single standardized form. As such, we have documented each data resource using R package documentation, allowing the metadata of each data product to be examined directly from R using the \texttt{help()} function or the question mark notation (e.g., \texttt{?viruses}).

\paragraph*{Data cleaning and taxonomic resolution}

The initial data structure was maintained from the original raw data files provided by David Onstad, principal maintainer of the EDWIP data resource \citep{onstad1997}. This includes files such as \texttt{new\_assoc}, as this was likely a test file containing pathogen species such as "wormy thing". We clean and augment this existing data source programmatically, with much of this code in the \texttt{insectDisease} package \texttt{vignette}.

\paragraph*{}
First, we resolve host and pathogen names using the R package \texttt{taxize}, using NCBI taxonomic backbone \citep{chamberlain2013}. While this seems like a small change, it both standardizes host and pathogen nomenclature, and catches any taxonomic changes that have occurred in the past couple decades. This includes the consideration of microsporidian parasites as fungi, not protozoans, a change affecting a large set of records in the EDWIP data. Second, we provide a vignette which combines the different host-pathogen data together to form an informative set of host-pathogen associations, including both both known associations and failed experimental infections, as delineated by the \texttt{interaction} column in the resulting \texttt{edwip.csv} file. Finally, processed data by the vignette is output in \texttt{.csv} format in the vignettes folder, but all data stored as R data objects (\texttt{.rda}) are also converted to csv and placed in the \texttt{csv} folder, allowing non-R users to still access the data in a stable format.

\subsection*{Case study: covariance among pathogen groups in parasite species richness }

Hosts that are infected by more pathogens of one type may also be more infected by pathogens of another type, mediated by host life history traits, metabolic demands, geographic distribution, and intensity of scientific study \citep{dallas2021}. We explore this in the EDWIP data by measuring the number of known positive associations of each of the pathogen groups for each insect host species, visualizing the relationship between the number of pathogens per insect host as a correlation matrix (Figure \ref{fig:corPlot}). We find very little evidence that pathogen groups have positive covariance, which would be expected if host species traits drove infection process across pathogen groups in the same manner. The failure to detect positive relationships, and indeed some negative relationships appearing, could be a signal of the targeted nature of data collection, as many insect host species were selected to study due to their potential as a crop pest, and many pathogens were selected to study based on their potential use as biocontrol or perhaps for their ease of culture.

\paragraph*{}
This potential sampling bias among insect host species would be clear if there were a positive relationship between the number of positive interactions to the number of negative interactions for a host species, as it would indicate that insect host species with lots of known interactions also tended to appear in many studies and have some negative interactions as well. We find evidence for a significantly negative relationship based on a Spearman's rank correlation ($\rho$ = -0.1, p $<$ 0.0001).

\subsection*{Concluding comments}

Ecological data -- while growing in availability, size, and stability -- are still a resource that should not be allowed to fade out of existence. The EDWIP data provided to the authors were in a proprietary format ('Claris FileMaker Pro 5') that was already over 10 major versions behind. With limited inter-version operability (e.g., \texttt{.fmp5} files cannot be opened in more recent versions of the software, or require multiple conversion steps), these data seemed as if headed towards obsolescence. The \texttt{insectDisease} package ensures that these data will be available to the broadest set of researchers, be bound to relevant metadata, and be properly versioned. By hosting the data openly, we welcome contributions from researchers interested in augmenting the data or building off the existing resource.

\subsection*{Data accessibility}

The \texttt{insectDisease} R package is currently available on GitHub https://github.com/viralemergence/insectDisease, with '.csv' files in the csv directory for long-term data stability.

Finally, the data are periodically archived at major version changes via Zenodo ??????????????????

\clearpage
\bibliography{var}
\bibliographystyle{ecography}
\clearpage

\clearpage

%----------------
\section*{Figures}

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.85\textwidth]{Figures/bubble.pdf}
    \caption{Bubble plot, where points are proportional to the total number of negative (on the left) and positive (on the right) host-pathogen interactions for each pathogen group ($y$-axis). Numeric columns correspond to the number of unique host species, pathogen species, and interactions for each pathogen group. }
    \label{fig:bubble}
  \end{center}
\end{figure}

\clearpage

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{Figures/corPlot.pdf}
    \caption{Correlations between each pathogen group in terms of pathogen richness of insect host species, where color corresponds to Spearman's rank correlation values (provided in the lower diagonal matrix). Fungal and protozoan pathogens seemed to negatively covary, as did viruses and protozoans. Understanding to what extent this is driven by sampling effects or insect host ecology is an outstanding research question that these could be used to begin addressing.}
    \label{fig:corPlot}
  \end{center}
\end{figure}

\end{document}

### 2021-08-11 R/InsectDisease-package.r

@@ -12,7 +12,7 @@
 #' *When using this database, cite this reference*
 #'
 #' Braxton, S. M., et al. "Description and analysis of two internet-based databases of insect
-#' pathogens: EDWIP and VIDIL." Journal of invertebrate pathology 83.3 (2003): 185-195.
+#' pathogens: EDWIP and VIDIL." Journal of Invertebrate Pathology 83.3 (2003): 185-195.
 #' 
 #' Also, this database was originally created by the following people, to whom we are indebted:
 #'

### 2021-08-11 R/hostTaxonomy.R
trying to get vigneette to work

@@ -1,0 +1,23 @@
+#' Cached version of host taxonomy (see vignette for code to build from scratch)
+#'
+#' 
+#' 
+#' 
+#'
+#' \itemize{
+#'   \item HostOriginal original host name from EDWIP
+#'   \item HostTaxID NCBI taxonomic ID
+#'   \item HostNCBIResolved is the host found in NCBI
+#'   \item HostGenus host genus
+#'   \item HostFamily host family
+#'   \item HostOrder host order
+#'   \item HostClass  host class
+#'  }
+#'
+#' @docType data
+#' @keywords datasets
+#' @name hostTaxonomy
+#' @usage data(hostTaxonomy)
+#' @format A data.frame 
+
+"hostTaxonomy"

### 2021-08-11 R/pathTaxonomy.R

@@ -1,0 +1,20 @@
+#' Cached version of pathogen taxonomy (see vignette for code to build from scratch)
+#' 
+#'
+#' \itemize{
+#'   \item PathTaxID NCBI taxonomic ID
+#'   \item PathNCBIResolved is the pathogen found in NCBI
+#'   \item PathGenus pathogen genus
+#'   \item PathFamily pathogen family
+#'   \item PathOrder pathogen order
+#'   \item PathClass  pathogen class
+#'   \item PathKingdom  pathogen class
+#'  }
+#'
+#' @docType data
+#' @keywords datasets
+#' @name pathTaxonomy
+#' @usage data(pathTaxonomy)
+#' @format A data.frame 
+
+"pathTaxonomy"

### 2021-08-16 R/getNCBI.R
travis ci to github actions migration

@@ -1,0 +1,60 @@
+#' Get NCBI taxonomy data
+#'
+#' Process through a vector of species names to obtain taxonomic data
+#' 
+#' @param species a vector of species names
+#' @param host (boolean) affects column naming (nice to keep host and pathogen separate)
+#'
+#' @return a data.frame with nrow == length(species) 
+#' @name getNCBI
+#' @usage getNCBI('Peromyscus')
+
+
+getNCBI <- function(species, host=TRUE){ 
+  names.orig <- species
+  species <- str_replace(species, " cf\\.","")
+  species <- str_replace(species, " sp\\.","")
+  species <- str_replace(species, " gen\\.","")
+  u <- get_uid(species, 
+    rank_filter = c("subspecies", "species", "genus", "family", "order", "class"), 
+    division_filter = "vertebrates", ask = FALSE)
+  c <- classification(u)
+  n <- !is.na(u)
+  attributes(u) <- NULL
+  s <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="species")]], error = function(e) {NA})}), use.names = FALSE)
+  g <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="genus")]], error = function(e) {NA})}), use.names = FALSE)
+  f <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="family")]], error = function(e) {NA})}), use.names = FALSE)
+  o <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="order")]], error = function(e) {NA})}), use.names = FALSE)
+  c2 <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="class")]], error = function(e) {NA})}), use.names = FALSE)
+  k <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="kingdom")]], error = function(e) {NA})}), use.names = FALSE)  
+  levels <- c("species", "genus", "family", "order", "class")
+  u <- unlist(lapply(c, function(x){tryCatch(last(na.omit(x[x$rank %in% levels,'id'])), 
+    error = function(e) {NA})}), use.names = FALSE)
+  
+  if(host){  
+    ret <- data.frame(HostOriginal = names.orig,
+             HostTaxID = u,
+             HostNCBIResolved = n, 
+             Host = s,
+             HostGenus = g,
[…]

### 2021-08-16 R/zzz.R

@@ -1,0 +1,86 @@
+#' new_asso
+#'
+#'
+#' \itemize{
+#'   \item ERNnew
+#'   \item HostSpecies
+#'   \item HostOrder
+#'   \item HostFamily
+#'   \item HostHabitat
+#'   \item HostFood
+#'   \item HostGenYr
+#'   \item PathSpecies
+#'   \item PathGroup
+#'   \item PathHighTaxon
+#'   \item PathLowTaxon
+#'   \item StageInf
+#'   \item TissueInfected
+#'   \item Field
+#'   \item Country
+#'   \item IntermediateHost
+#'   \item Citation
+#'   \item MoreInfo
+#'   \item Who
+#'   \item CreationDate
+#'   \item ModificationDate
+#'   \item StainFCB
+#'  }
+#'
+#' @docType data
+#' @keywords datasets
+#' @name new_asso
+#' @usage data(new_asso)
+#' @format A data.frame of not good records
+
+
+
+
+#' newnema
+#'
+#'
+#' \itemize{
+#'   \item ERNnem. EDWIP record number
+#'   \item Host
+#'   \item Nematode
+#'   \item NemaOrder
+#'   \item NemaFamily
+#'   \item NemaStrain
+#'   \item StageInfected
+#'   \item TissueInfected
+#'   \item FieldOrLab
+#'   \item Country
+#'   \item SoilType
+#'   \item AssociatedBacterium
+#'   \item IntermediateHost
+#'   \item CreationDate
+#'   \item ModificationDate
+#'  }
+#'
+#' @docType data
+#' @keywords datasets
+#' @name new_asso
+#' @usage data(newnema)
+#' @format A data.frame 
+
+
+
+
+
+#' noassref
+#'
+#'
+#' \itemize{
+#'   \item RefCode. Index of reference 
+#'   \item ERNntr. EDWIP record number
+#'   \item Host. Host species
+#'   \item Pathogen. Pathogen species
+#'   \item Citation. Citation for host-nematode record
+#'  }
+#'
+#' @docType data
+#' @keywords datasets
+#' @name noassref
+#' @usage data(noassref)
+#' @format A data.frame 
+
+

### 2021-08-16 R/getNCBI.R
travis ci to github actions migration

@@ -7,6 +7,7 @@
 #'
 #' @return a data.frame with nrow == length(species) 
 #' @name getNCBI
+#' @export
 #' @usage getNCBI('Peromyscus')
 
 

### 2021-08-16 R/zzz.R

@@ -32,7 +32,7 @@
 #' @usage data(new_asso)
 #' @format A data.frame of not good records
 
-
+"new_asso"
 
 
 #' newnema
@@ -62,7 +62,7 @@
 #' @usage data(newnema)
 #' @format A data.frame 
 
-
+"newnema"
 
 
 
@@ -83,4 +83,6 @@
 #' @usage data(noassref)
 #' @format A data.frame 
 
+
+"noassref"
 

### 2021-08-16 R/getNCBI.R
fixed some warnings. non-ascii char still present in many files

@@ -7,19 +7,29 @@
 #'
 #' @return a data.frame with nrow == length(species) 
 #' @name getNCBI
+#' @importFrom taxize classification get_uid
+#' @importFrom stats na.omit 
+#' @importFrom utils tail
+#' @importFrom dplyr mutate
 #' @export
-#' @usage getNCBI('Peromyscus')
+
 
 
 getNCBI <- function(species, host=TRUE){ 
+
+  mutate_cond <- function(.data, condition, ..., envir = parent.frame()) {
+    condition <- eval(substitute(condition), .data, envir)
+    .data[condition, ] <- dplyr::mutate(.data[condition, ], ...)
+    .data
+  }
   names.orig <- species
-  species <- str_replace(species, " cf\\.","")
-  species <- str_replace(species, " sp\\.","")
-  species <- str_replace(species, " gen\\.","")
-  u <- get_uid(species, 
+  species <- gsub(" cf\\.","", species)
+  species <- gsub(" sp\\.","", species)
+  species <- gsub(" gen\\.","",species)
+  u <- taxize::get_uid(species, 
     rank_filter = c("subspecies", "species", "genus", "family", "order", "class"), 
     division_filter = "vertebrates", ask = FALSE)
-  c <- classification(u)
+  c <- taxize::classification(u)
   n <- !is.na(u)
   attributes(u) <- NULL
   s <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="species")]], error = function(e) {NA})}), use.names = FALSE)
@@ -29,18 +39,20 @@ getNCBI <- function(species, host=TRUE){
   c2 <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="class")]], error = function(e) {NA})}), use.names = FALSE)
   k <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="kingdom")]], error = function(e) {NA})}), use.names = FALSE)  
   levels <- c("species", "genus", "family", "order", "class")
-  u <- unlist(lapply(c, function(x){tryCatch(last(na.omit(x[x$rank %in% levels,'id'])), 
-    error = function(e) {NA})}), use.names = FALSE)
+  u <- unlist(lapply(c, function(x){
+    tryCatch(tail(na.omit(x[x$rank %in% levels,'id']),1), 
+      error = function(e) {NA})}), use.names = FALSE)
   
   if(host){  
[…]

### 2021-08-17 R/hostTaxonomy.R
fixed garbo non-ascii char issues

@@ -1,8 +1,7 @@
 #' Cached version of host taxonomy (see vignette for code to build from scratch)
 #'
 #' 
-#' 
-#' 
+#' @description Host taxonomic cached data from NCBI (see vignette for code to re-make)
 #'
 #' \itemize{
 #'   \item HostOriginal original host name from EDWIP

### 2021-08-26 R/assocref.R
updated with tweaks to the raw data since it was kinda iffy

@@ -7,8 +7,19 @@
 #'   \item ERNnvp. EDWIP record number 
 #'   \item refCode. Reference code
 #'   \item Reference. Actual citation
-#'   \item Host. Host species
-#'   \item Pathogen. Pathogen species
+#'   \item HostSpecies. Host species
+#'   \item PathogenSpecies. Pathogen species
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/getNCBI.R

@@ -28,8 +28,8 @@ getNCBI <- function(species, host=TRUE){
   species <- gsub(" gen\\.","",species)
   u <- taxize::get_uid(species, 
     rank_filter = c("subspecies", "species", "genus", "family", "order", "class"), 
-    division_filter = "vertebrates", ask = FALSE)
-  c <- taxize::classification(u)
+      ask = FALSE)
+  c <- taxize::classification(u, batch_size=50, max_tries=500)
   n <- !is.na(u)
   attributes(u) <- NULL
   s <- unlist(lapply(c, function(x){tryCatch(x$name[[which(x$rank=="species")]], error = function(e) {NA})}), use.names = FALSE)
@@ -44,29 +44,24 @@ getNCBI <- function(species, host=TRUE){
       error = function(e) {NA})}), use.names = FALSE)
   
   if(host){  
-    ret <- data.frame(HostOriginal = names.orig,
+    ret <- data.frame(
+      HostSpecies = names.orig,
       HostTaxID = u,
-      HostNCBIResolved = n, 
-      Host = s,
       HostGenus = g,
       HostFamily = f,
       HostOrder = o, 
       HostClass = c2) 
-    ret <- mutate_cond(ret, ret$HostNCBIResolved == FALSE, Host = ret$HostOriginal) 
     ret$Host <- NULL
   }
   if(host==FALSE){
     ret <- data.frame(
-      PathOriginal=names.orig, 
+      PathogenSpecies=names.orig, 
       PathTaxID = u,
-      PathNCBIResolved = n, 
       PathGenus = g,
       PathFamily = f,
       PathOrder = o, 
       PathClass = c2,
       PathKingdom= k) 
-    ret <- mutate_cond(ret, ret$PathNCBIResolved == FALSE, Path = ret$PathOriginal) 
-    ret$PathOriginal <- NULL
   }
   return(ret)
 }

### 2021-08-26 R/hostTaxonomy.R

@@ -4,9 +4,8 @@
 #' @description Host taxonomic cached data from NCBI (see vignette for code to re-make)
 #'
 #' \itemize{
-#'   \item HostOriginal original host name from EDWIP
+#'   \item HostSpecies original host name from EDWIP
 #'   \item HostTaxID NCBI taxonomic ID
-#'   \item HostNCBIResolved is the host found in NCBI
 #'   \item HostGenus host genus
 #'   \item HostFamily host family
 #'   \item HostOrder host order

### 2021-08-26 R/hosts.R

@@ -27,6 +27,11 @@
 #'   \item CommonNameOther. Other common names?
 #'   \item Complete. Is this record complete?
 #'   \item AdditionalReferences. Additional reference indices. 
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/negative.R

@@ -8,17 +8,26 @@
 #' 
 #' \itemize{
 #'   \item ERNntr. EDWIP record number
-#'   \item Pathogen. Pathogen species
+#'   \item PathogenSpecies. Pathogen species
 #'   \item DateEntered. Date of initial data entry
 #'   \item DateModified. Modification date of entry
 #'   \item LogMaxDose. Dosage, in many different units
 #'   \item HostStageTested. Host stage exposed to pathogen (e.g. Larvae, Nymph, Adult)
-#'   \item Host. Host species examined
+#'   \item HostSpecies. Host species examined
 #'   \item Group. Pathogen group (e.g. viruses) 
 #'   \item HighTaxon. General classification of pathogen (e.g. DNA virus)
 #'   \item LowTaxon. More specific classification of pathogen (e.g. Baculoviridae)
-#'   \item Order. Host taxonomic order.
-#'   \item Family. Host taxonomic family.
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/nemaref.R

@@ -8,8 +8,19 @@
 #'   \item ERNnem. EDWIP record number
 #'   \item refCode. Index of reference obtained from `nematode` data frame
 #'   \item Reference. Citation for host-nematode record
-#'   \item Host. Host species
-#'   \item Nematode. Nematode parasite species
+#'   \item HostSpecies. Host species
+#'   \item PathogenSpecies. Nematode parasite species
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/nematode.R

@@ -5,11 +5,9 @@
 #' 
 #' \itemize{
 #'   \item ERNnem
-#'   \item Host. Host species
-#'   \item Nematode. Nematode parasite species
-#'   \item NematodeOrder. Nematode parasite order
-#'   \item NematodeFamily. Nematode parasite family
-#'   \item NematodeStrain. Nematode parasite strain
+#'   \item HostSpecies. Host species
+#'   \item PathogenSpecies. Nematode parasite species
+#'   \item PathogenStrain. Nematode parasite strain
 #'   \item StageInfected. Host stage infected
 #'   \item TissueInfected. Host tissue infected
 #'   \item FieldOrLab. Was this a field or lab tested association?
@@ -19,6 +17,17 @@
 #'   \item IntermediateHost. Is there an intermediate host present?
 #'   \item CreationDate. Date of initial data entry
 #'   \item ModificationDate. Modification date of entry
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/nvpassoc.R

@@ -7,13 +7,9 @@
 #' 
 #' \itemize{
 #'   \item ERNnvp. EDWIP record number
-#'   \item Pathogen. Pathogen species
+#'   \item PathogenSpecies. Pathogen species
 #'   \item Group. Pathogen group (e.g. Protozoa)
-#'   \item HighTaxon. General classification of pathogen (e.g. Microspora)
-#'   \item LowTaxon. More specific classification of pathogen (e.g. Microsporea)
-#'   \item Host. Host species examined
-#'   \item Order. Host taxonomic order.
-#'   \item Family. Host taxonomic family.
+#'   \item HostSpecies. Host species examined
 #'   \item HostStageTested. Host stage exposed to pathogen (e.g. Larvae, Nymph, Adult)
 #'   \item HostTissueInfected. Host tissue infected
 #'   \item FieldOrLab. Was this a field or lab tested association?
@@ -22,6 +18,17 @@
 #'   \item DateEntered. Date of initial data entry
 #'   \item DateModified. Modification date of entry
 #'   \item BiogeographicRegion. Biogeographic region (or some combination thereof)
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/pathogen.R

@@ -5,9 +5,14 @@
 #' \itemize{
 #'   \item PathogenSpecies. Pathogen species.
 #'   \item Group. Pathogen group (e.g. Protozoa)
-#'   \item HighTaxon. General classification of pathogen (e.g. Microspora)
-#'   \item LowTaxon. More specific classification of pathogen (e.g. Microsporea)
 #'   \item AdditionalNotes. Some additional notes.
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
+
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/viraref.R

@@ -8,8 +8,20 @@
 #'   \item RefCode. Index of reference obtained from `nematode` data frame
 #'   \item Citation. Reference.
 #'   \item ERNv. EDWIP record number
-#'   \item Host. Host species
-#'   \item Virus. Virus name
+#'   \item HostSpecies. Host species
+#'   \item PathogenSpecies. Virus name
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
+
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/viruses.R

@@ -5,9 +5,10 @@
 #' 
 #' \itemize{
 #'   \item ERNv. EDWIP record number
-#'   \item Host. Host species
-#'   \item HighTaxon. General classification of pathogen (e.g. Microspora)
-#'   \item LowTaxon. More specific classification of pathogen (e.g. Microsporea)
+#'   \item HostSpecies. Host species
+#'   \item VirusType. DNA or RNA virus
+#'   \item Group. Viruses.
+#'   \item PathogenSpecies. Viral family 
 #'   \item Virus. Virus identity
 #'   \item HostStageInfected. Host stage infected
 #'   \item HostTissueInfected. Host tissue infected
@@ -18,6 +19,17 @@
 #'   \item ModificationDate. Modification date of entry (wrong)
 #'   \item ProvinceA. Canadian provinces where host-virus interaction occurs
 #'   \item PathogenValue. Is there value to the pathogen (can it be used as a control agent?)
+#'   \item HostTaxID. Host NCBI ID number
+#'   \item HostGenus. Host genus
+#'   \item HostFamily. Host family
+#'   \item HostOrder. Host order
+#'   \item HostClass. Host class
+#'   \item PathTaxID. Pathogen NCBI ID number
+#'   \item PathGenus. Pathogen genus
+#'   \item PathFamily. Pathogen family
+#'   \item PathOrder. Pathogen order
+#'   \item PathClass. Pathogen class
+#'   \item PathKingdom. Pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2021-08-26 R/zzz.R

@@ -1,6 +1,9 @@
 #' new_asso
 #'
-#'
+#' These data are included because they were originally part of the EDWIP data. 
+#' We caution the user to not use these data, as we do not believe the records 
+#' are correct. 
+#' 
 #' \itemize{
 #'   \item ERNnew
 #'   \item HostSpecies
@@ -37,6 +40,9 @@
 
 #' newnema
 #'
+#' These data are included because they were originally part of the EDWIP data. 
+#' These data are identical to the `nematode` data, as far as we can tell.
+#' 
 #'
 #' \itemize{
 #'   \item ERNnem. EDWIP record number
@@ -68,13 +74,16 @@
 
 #' noassref
 #'
+#' These data are included because they were originally part of the EDWIP data. 
+#' We caution the user in using these data, as they do not have associated 
+#' metadata that the other records have.
 #'
 #' \itemize{
 #'   \item RefCode. Index of reference 
 #'   \item ERNntr. EDWIP record number
 #'   \item Host. Host species
 #'   \item Pathogen. Pathogen species
-#'   \item Citation. Citation for host-nematode record
+#'   \item Citation. Citation for host-pathogen record
 #'  }
 #'
 #' @docType data

### 2021-08-31 manuscript/manuscript.tex
updates to vignette

@@
 \paragraph*{}
-Data are also available on the insect host species themselves (e.g., \texttt{?hosts} or \texttt{data(hosts)}).
+Data are also available on the insect host species themselves (e.g., \texttt{data(hosts)}).
 These data contain some information on Canadian province where the host is found (\texttt{ProvinceI} column), what it eats (\texttt{Food} column), and what type of habitat it is found in (\texttt{Habitat} column).
@@
 
-The initial data structure was maintained from the original raw data files provided by David Onstad, principal maintainer of the EDWIP data resource \citep{onstad1997}.
+We attempted to maintain as much of the original data structure from the raw data files provided by David Onstad, principal maintainer of the EDWIP data resource \citep{onstad1997}.
 This includes files such as \texttt{new\_assoc}, as this was likely a test file containing pathogen species such as "wormy thing".
+We provide comments in the documentation about which data sources are perhaps more trustworthy.
 We clean and augment this existing data source programmatically, with much of this code in the \texttt{insectDisease} package \texttt{vignette}.
@@
 \paragraph*{}
-First, we resolve host and pathogen names using the R package \texttt{taxize}, using NCBI taxonomic backbone \citep{chamberlain2013}.
-While this seems like a small change, it both standardizes host and pathogen nomenclature, and catches any taxonomic changes that have occurred in the past couple decades.
+The first, and perhaps most important augmentation, is the resolution of host and pathogen taxonomic information.
+We achieved this by using the R package \texttt{taxize}, specifically the NCBI taxonomic backbone \citep{chamberlain2013}.
+Cached versions of host and pathogen taxonomic information are provided (\texttt{data(hostTaxonomy)} and \texttt{data(pathTaxonomy)}), and the $R$ code to generate these taxonomic backbones and clean the data are provided in the package vignette.
+This taxonomic backbone serves to both standardize host and pathogen nomenclature, while also correcting any taxonomic changes that have occurred in the past couple decades.
 This includes the consideration of microsporidian parasites as fungi, not protozoans, a change affecting a large set of records in the EDWIP data.
-Second, we provide a vignette which combines the different host-pathogen data together to form an informative set of host-pathogen associations, including both both known associations and failed experimental infections, as delineated by the \texttt{interaction} column in the resulting \texttt{edwip.csv} file.
-Finally, processed data by the vignette is output in \texttt{.csv} format in the vignettes folder, but all data stored as R data objects (\texttt{.rda}) are also converted to csv and placed in the \texttt{csv} folder, allowing non-R users to still access the data in a stable format.
+Second, we provide a vignette which transparently shows the steps to clean and augment the data resource, as well as reproduce figures from this manuscript.
+Further, the vignette includes code which generates host-pathogen associations for all the data combined into a single data resource, simplifying the data down to a small number of important columns.
+Finally, processed we opt to store raw data in the \texttt{csv} folder, which contains all raw data files in .csv format.
+These data are also provided as .rda files in the \texttt{data} folder.
+This allows non-$R$ users to access the csv-formatted data easily, and ensures long-term stability of the data, as \texttt{csv} is a stable text file format.
 

### 2021-12-07 R/hosts.R
updates based on reviewer comments

@@ -21,8 +21,8 @@
 #'     7 unique values
 #'   \item ModificationDate. Modification date of entry
 #'   \item InCanada. Citations for presence/absence of host in Canada. Numeric indices can be
-#'     related to the citations in the `citations.rda` data file. `Y` and `N` relate to presence and
-#'     absence, respectively.
+#'     related to the citations in the `citations.rda` data file. `Y` and `N` 
+#'     relate to presence and absence, respectively.
 #'   \item ChangeSpeciesTo. 
 #'   \item CommonNameOther. Other common names?
 #'   \item Complete. Is this record complete?

### 2021-12-07 data-raw/process.R

@@ -1,0 +1,259 @@
+
+#' Programmatically adjust pathogen "Group" column based on taxonomy
+#'
+#' @param dat the data.frame containing a column "Group"
+#' @param tax the taxonomy data.frame from above
+#'
+#' @return dat, but with updated "Group" designations based on taxonomy
+
+fillTaxon <- function(edwip){
+    
+  protzC <- c('Kinetoplastea')
+  bactC <- c('Gammaproteobacteria', 'Bacilli')
+  
+  fungz <- c('Unikaryon ixodis', 'Tompetchia webberi', 
+    'Sphaerostilbe auranticola', 'Erynia sepulchralis',
+    'Furia vomitoriae', 'Hymenostilbe sp.', 'Hymenostilbe spp.', 
+    'Aquamortierella elegans', "Sorosporella sp.", "Coreomycetopsis oedipus",
+    "Termitaria coronata", "Herpomyces paranensis", "Herpomyces tricuspidatus", 
+    "Herpomyces arietinus", "Herpomyces nyctoborae", "Strongwellsea sp. nov. 1", 
+    "Strongwellsea sp. nov. 2", "Strongwellsea sp. nov. 3", 
+    "Strongwellsea sp. nov. 6", "Strongwellsea sp. nov. 5", 
+    "Strongwellsea sp. nov. 4", "Eryniopsis lampyridarum", 
+    "Entomophthora erupta", "Erynia erinacea", "Erynia spp.",
+    "Thaxterosporium turbinatum", "Diheterospora sp.", "Neozygites cf floridana", 
+    "Coelomomyces sp.", "Pandora gammae", "Myiophagus sp.", 
+    "Septobasidium spp.", "Neozygites cucumeriformis", 
+    "Synnematium sp.", "Fusarium oxysporum var. orthoceras", 
+    "Fusarium sp. near solani","Apergillus sp.","Leptolegnia sp.", 
+    "Zoophthora curvispora")
+
+  virz <- c("Thymelicus lineola NPV", 
+    "Carposina niponensis GV", "Colias eurytheme NPV", 
+    "Celerio euphorbiae NPV", "Aedes taeniorhynchus IV (RMIV)", 
+    "Bombyx mori CPV", "Dendrolimus spectabilis CPV")
+
+  protz <- c('Gregarina guatemalensis', 'Hoplorhynchus polyhamatus', 
[…]

### 2021-12-07 manuscript/webVersion/bin/db_setup.py

@@ -1,0 +1,52 @@
+import sqlite3
+import csv
+
+def escape(s):
+    return s.replace('"','""').replace("'","''").strip()
+
+db=sqlite3.connect("data/db.sqlite")
+c=db.cursor()
+
+c.execute('CREATE TABLE Main (ID INTEGER PRIMARY KEY, Host TEXT, ParasiteGroup TEXT, HighTaxon TEXT, LowTaxon TEXT, Pathogen TEXT, Infect INTEGER, Stage TEXT, Tissue, TEXT, FieldOrLab TEXT, Country TEXT)')
+with open('data/csv/negative.csv', newline='', encoding='iso-8859-1') as csvfile:
+    reader = csv.DictReader(csvfile)
+    for row in reader:
+        c.execute(f"INSERT INTO Main (Host, ParasiteGroup, HighTaxon, LowTaxon, Pathogen, Infect, Stage, Tissue, FieldOrLab, Country) VALUES (\"{escape(row['Host'])}\", \"{escape(row['Group'])}\", \"{escape(row['HighTaxon'])}\", \"{escape(row['LowTaxon'])}\", \"{escape(row['Pathogen'])}\", \"0\", \"{escape(row['HostStageTested'])}\", \"NA\", \"Lab\", \"NA\")")
+
+with open('data/csv/nematode.csv', newline='', encoding='iso-8859-1') as csvfile:
+    reader = csv.DictReader(csvfile)
+    for row in reader:
+        c.execute(f"INSERT INTO Main (Host, ParasiteGroup, HighTaxon, LowTaxon, Pathogen, Infect, Stage, Tissue, FieldOrLab, Country) VALUES (\"{escape(row['Host'])}\", \"nematode\", \"{escape(row['NematodeOrder'])}\", \"{escape(row['NematodeFamily'])}\", \"{escape(row['Nematode'])}\", \"1\", \"{escape(row['StageInfected'])}\", \"{escape(row['TissueInfected'])}\", \"{escape(row['FieldOrLab'])}\", \"{escape(row['Country'])}\")")
+
+with open('data/csv/nvpassoc.csv', newline='', encoding='iso-8859-1') as csvfile:
+    reader = csv.DictReader(csvfile)
+    for row in reader:
[…]

### 2021-12-07 manuscript/webVersion/bin/form_values.py

@@ -1,0 +1,55 @@
+def query_list(field, db, table, args=()):
+    db.row_factory = lambda c, r : r[0]
+    cur = db.execute(f'SELECT {field} from {table}', args)
+    rows = [row for row in cur.fetchall() if row != 'no']
+    cur.close()
+    return sorted(list(set(rows)))
+
+def query_colnames(table, db, exclude=[], args=()):
+    db.row_factory = lambda c, r : r[0]
+    cur = db.execute(f'SELECT name FROM PRAGMA_TABLE_INFO(\'{table}\')')
+    cols = [col for col in cur.fetchall() if col not in exclude]
+    cur.close()
+    return sorted(list(set(cols)))
+
+def dropdowns(db, table='main'):
+    return {
+        'host_binomial'     : query_list('HostBinomial', db, 'Hosts'),
+        'host_order'        : query_list('HostOrder', db, 'Hosts'),
+        'host_family'       : query_list('HostFamily', db, 'Hosts'),
+        'host_genus'        : query_list('HostGenus', db, 'Hosts'),
+        'pathogen_name'     : query_list('Pathogen', db, 'Main'),
+        'pathogen_group'    : query_list('ParasiteGroup', db, 'Main'),
+        'status' : ['Both', 'Positive', 'Negative']
+    }
+
+def boxes():
+    return {
+        'host_fields' : ["Host Binomial (genus and species)"],
+        'parasite_fields' : ["Parasite Group (virus, bacteria, fungus, nematode, ect.)", "High Taxon (higher taxonomy)", "Low Taxon (lower taxonomy)", "Pathogen (genus and species)"],
+        'infection_fields' : ["Infect (infection status)", "Stage (host stage infected/ tested)", "Tissue (tissue infected/ tested)", "Field Or Lab (field or experimental observation)"],
+        'other_fields' : ["Country (country of host population)", "Host Order/ Taxon (higher host taxonomy)", "Host Family (lower host taxonomy)",
+            "Synonyms (alternate host scientific names)", "Habitat (host habitat)", "Food (host diet)",
+            "genYr (host generations per year)", "Common Name (host common name)", "Insect Status (pest or non-pest)"]
+    }
+
+columnMap = {
+    "hostbinomial(genusandspecies)"                         : "HostBinomial",
+    "parasitegroup(virus,bacteria,fungus,nematode,ect.)"    : "ParasiteGroup",
[…]

### 2021-12-07 manuscript/webVersion/bin/gunicorn.sh

@@ -1,0 +1,3 @@
+#!/bin/sh
+
+gunicorn -w 2 --threads 2 -b 0.0.0.0:5000 wsgi:app

### 2021-12-07 manuscript/webVersion/bin/query.py

@@ -1,0 +1,36 @@
+def makeQuery(data):
+    infectionStatus = data['InfectionStatus']
+    del(data['InfectionStatus'])
+
+    fieldlist = []
+    joinHosts = False
+    for field in [data[field][0] for field in data if '_fields_' in field]:
+        if field in ['HostBinomial', 'HostOrder', 'HostFamily', 'Synonyms', 'Habitat', 'Food', 'genYr', 'CommonName', 'InsectStatus']:
+            fieldlist.append(f'Hosts.{field}')
+            joinHosts = True
+        else:
+            fieldlist.append(f'Main.{field}')
+
+    fieldString = ', '.join(fieldlist)
+    if not fieldString:
+        fieldString = 'NULL'
+
+    queryString = f'SELECT {fieldString} FROM Main '
+    #if joinHosts:
+    queryString += f'INNER JOIN Hosts ON Main.Host = Hosts.HostBinomial '
+    wheres = ' AND '.join(['( '+' OR '.join([f'({field} = \'{value}\')' for value in data[field]])+ ' )' for field in data if '_fields_' not in field])
+    if wheres:
+        queryString += 'WHERE ' + wheres
+    if infectionStatus == 'Positive':
+        queryString += ' AND ( Infect = 1 )'
+    if infectionStatus == 'Negative':
+        queryString += ' AND ( Infect = 0 )'
+    return queryString
+
+def query(data, db):
+    query = makeQuery(data)
+    #print(query)
+    cur = db.execute(query,{})
+    records = cur.fetchall()
+    cur.close()
+    return [list(x) for x in records]

### 2021-12-07 manuscript/webVersion/bin/serve.py

@@ -1,0 +1,65 @@
+import sqlite3
+from flask import Flask, request, render_template, g, make_response
+
+import sys
+sys.path.append('./bin')
+
+import form_values
+import query
+
+app = Flask(__name__, root_path='./')
+
+def enquote(x):
+    if type(x) == str:
+        if ',' in x:
+            return f'"{x}"'
+        else:
+            return x
+    else:
+        return str(x)
+
+def make_dicts(cursor, row):
+    return dict((cursor.description[idx][0], value)
+                for idx, value in enumerate(row))
+
+def get_db(factory=None):
+    db = getattr(g, '_database', None)
+    if db is None:
+        db = g._database = sqlite3.connect('data/db.sqlite')
+        if factory:
+            db.row_factory = factory
+    return db
+
+@app.errorhandler(404)
+def page_not_found(e):
+    return render_template('404.html'), 404
+
+@app.errorhandler(500)
+def internal_error(e):
+    return render_template('500.html'), 500
+
+@app.route("/", methods=['GET'])
+def formPage():
+    return render_template('welcome.html', active='welcome'), 200
+    return '', 200
+
+@app.route("/download", methods=['GET', 'POST'])
+def queryPage():
+    if request.method == 'GET':
+        return render_template('form.html', dropdowns=form_values.dropdowns(get_db()), boxes=form_values.boxes(), active='download'), 200
+    if request.method == 'POST':
+        data = request.form.to_dict(flat=False)
+        for field in data:
+            if '_fields_' in field:
+                data[field] = [form_values.columnMap.get(data[field][0])]
+        #return(data)
+        results = query.query(data, get_db())
+        colnames = ','.join([data[field][0] for field in data if '_fields_' in field])
+        resp=make_response(colnames + '\n' + '\n'.join([','.join([enquote(x) for x in result]) for result in results]))
+        resp.headers.set('Content-Type', 'text/csv')
+        resp.headers.set("Content-Disposition", "attachment;filename=EDWIP_data.csv")
+        return(resp)
+    return render_template('500.html'), 500
+
+if __name__ == '__main__':
+    app.run(host='0.0.0.0',threaded=True,port=7778)

### 2021-12-07 manuscript/webVersion/static/js/bundle.js

@@ -1,0 +1,79 @@
+(function(){function r(e,n,t){function o(i,f){if(!n[i]){if(!e[i]){var c="function"==typeof require&&require;if(!f&&c)return c(i,!0);if(u)return u(i,!0);var a=new Error("Cannot find module '"+i+"'");throw a.code="MODULE_NOT_FOUND",a}var p=n[i]={exports:{}};e[i][0].call(p.exports,function(r){var n=e[i][1][r];return o(n||r)},p,p.exports,r,e,n,t)}return n[i].exports}for(var u="function"==typeof require&&require,i=0;i<t.length;i++)o(t[i]);return o}return r})()({"/Users/tomlin/RCN/parasites/parasites/static/js/index.js":[function(require,module,exports){
+'use strict';
+
+$(document).ready(function () {
+    $('#id_continent').on('hidden.bs.select', function (e, clickedIndex, newValue, oldValue) {
+        var continentSelected = $(this).val();
+        // hide options form country menu
+        var indexesToHide = [];
+        $('#id_country option').each(function (index) {
+            var continent = $(this).val().split('.')[0];
+            if (continentSelected && continentSelected.indexOf(continent) < 0) {
+                // if the continent for a country is not fund in the selected continents
+                // the country is hidden from selected.
+                indexesToHide.push(index);
+            }
+        });
[…]

### 2021-12-07 manuscript/webVersion/static/js/index.js

@@ -1,0 +1,73 @@
+$(document).ready(function() {
+    $('#id_continent').on('hidden.bs.select', function (e, clickedIndex, newValue, oldValue) {
+        var continentSelected = $(this).val();
+        // hide options form country menu
+        var indexesToHide = [];
+        $('#id_country option').each(function (index) {
+            var continent = $(this).val().split('.')[0];
+            if (continentSelected && continentSelected.indexOf(continent) < 0) {
+                // if the continent for a country is not fund in the selected continents
+                // the country is hidden from selected.
+                indexesToHide.push(index);
+            }
+        });
+        // go over widget options and hide as needed.
+        var $menu = $('#id_country').siblings('.dropdown-menu')[0];
+        $($menu).find('li').each(function(index) {
+           if (indexesToHide.indexOf(index) > -1) {
+               // this element needs to be hidden
+               $(this).addClass('hidden');
+               $(this).removeClass('selected');
+           } else {
+               // this element needs to be visible
+               $(this).removeClass('hidden');
+           }
+        });
+        // unselected previously selected values
+        var selectedCountries = $('#id_country').val(),
+            validSelectedCountries = [];
+        if (selectedCountries && selectedCountries.length) {
+            selectedCountries.forEach(function(item) {
+                var continent = item.split('.')[0];
+                if (continentSelected && continentSelected.indexOf(continent) > -1) {
+                    // country is in a selected continent so it is valid
+                    validSelectedCountries.push(item);
+                }
+            });
+            // only valid countries are set
+            $('#id_country').val(validSelectedCountries);
+            // check that the selected label is correct
+            var button = $('#id_country').siblings('button')[0],
+                selectWidgetLabel = $(button).find('.filter-option')[0],
+                currentSelection = $(selectWidgetLabel).text(),
+                labelStillValid;
+            if (validSelectedCountries) {
+                var numValidSelectedCountries = validSelectedCountries.length;
+                if (numValidSelectedCountries == 1) {
+                    // since there is only one valid country, let's display the country name
+                    $(button).attr('title', validSelectedCountries[0]);
+                    $(selectWidgetLabel).text(validSelectedCountries[0]);
+                } else {
+                    // display the number of items selected
+                    var labelMsg = numValidSelectedCountries + ' items selected';
+                    $(button).attr('title', labelMsg);
+                    $(selectWidgetLabel).text(labelMsg);
[…]

### 2021-12-07 manuscript/webVersion/static/libs/modernizr.js

@@ -1,0 +1,1440 @@
+/*!
+ * modernizr v3.11.1
+ * Build https://modernizr.com/download?-addtest-printshiv-setclasses-testprop-dontmin
+ *
+ * Copyright (c)
+ *  Faruk Ates
+ *  Paul Irish
+ *  Alex Sexton
+ *  Ryan Seddon
+ *  Patrick Kettner
+ *  Stu Cox
+ *  Richard Herrera
+ *  Veeck
+
+ * MIT License
+ */
+
+/*
+ * Modernizr tests which native CSS3 and HTML5 features are available in the
+ * current UA and makes the results available to you in two ways: as properties on
+ * a global `Modernizr` object, and as classes on the `<html>` element. This
+ * information allows you to progressively enhance your pages with a granular level
+ * of control over the experience.
+*/
+
+;(function(scriptGlobalObject, window, document, undefined){
+
+  var tests = [];
+  
+
+  /**
+   * ModernizrProto is the constructor for Modernizr
+   *
+   * @class
+   * @access public
+   */
+  var ModernizrProto = {
+    // The current version, dummy
+    _version: '3.11.1',
+
+    // Any settings that don't work as separate modules
+    // can go in here as configuration.
+    _config: {
+      'classPrefix': '',
+      'enableClasses': true,
+      'enableJSClass': true,
+      'usePrefixes': true
+    },
+
+    // Queue of tests
+    _q: [],
+
+    // Stub these for people who are listening
+    on: function(test, cb) {
+      // I don't really think people should do this, but we can
+      // safe guard it a bit.
+      // -- NOTE:: this gets WAY overridden in src/addTest for actual async tests.
+      // This is in case people listen to synchronous tests. I would leave it out,
+      // but the code to *disallow* sync tests in the real version of this
+      // function is actually larger than this.
+      var self = this;
+      setTimeout(function() {
+        cb(self[test]);
+      }, 0);
+    },
+
+    addTest: function(name, fn, options) {
+      tests.push({name: name, fn: fn, options: options});
+    },
+
+    addAsyncTest: function(fn) {
+      tests.push({name: null, fn: fn});
+    }
+  };
+
+  
+
+  // Fake some of Object.create so we can force non test results to be non "own" properties.
+  var Modernizr = function() {};
+  Modernizr.prototype = ModernizrProto;
+
+  // Leak modernizr globally when you `require` it rather than force it here.
[…]

### 2021-12-07 manuscript/webVersion/wsgi.py

@@ -1,0 +1,1 @@
+from bin.serve import app

### 2021-12-08 R/getNCBI.R
updates based on reviewer comments

@@ -10,7 +10,7 @@
 #' @importFrom taxize classification get_uid
 #' @importFrom stats na.omit 
 #' @importFrom utils tail
-#' @importFrom dplyr mutate
+#' @importFrom plyr mutate
 #' @export
 
 
@@ -19,7 +19,7 @@ getNCBI <- function(species, host=TRUE){
 
   mutate_cond <- function(.data, condition, ..., envir = parent.frame()) {
     condition <- eval(substitute(condition), .data, envir)
-    .data[condition, ] <- dplyr::mutate(.data[condition, ], ...)
+    .data[condition, ] <- plyr::mutate(.data[condition, ], ...)
     .data
   }
   names.orig <- species

### 2021-12-08 vignettes/dataManip.R

@@ -1,0 +1,8 @@
+## ---- results='hide'----------------------------------------------------------
+
+library(insectDisease)
+library(plyr)
+library(dplyr)
+library(corrplot)
+knitr::opts_chunk$set(fig.width=6, fig.height=6) 
+

### 2021-12-08 manuscript/webVersion/bin/db_setup.py
updates based on reviewer comments

@@ -1,52 +1,0 @@
-import sqlite3
-import csv
-
-def escape(s):
-    return s.replace('"','""').replace("'","''").strip()
-
-db=sqlite3.connect("data/db.sqlite")
-c=db.cursor()
-
-c.execute('CREATE TABLE Main (ID INTEGER PRIMARY KEY, Host TEXT, ParasiteGroup TEXT, HighTaxon TEXT, LowTaxon TEXT, Pathogen TEXT, Infect INTEGER, Stage TEXT, Tissue, TEXT, FieldOrLab TEXT, Country TEXT)')
-with open('data/csv/negative.csv', newline='', encoding='iso-8859-1') as csvfile:
-    reader = csv.DictReader(csvfile)
-    for row in reader:
-        c.execute(f"INSERT INTO Main (Host, ParasiteGroup, HighTaxon, LowTaxon, Pathogen, Infect, Stage, Tissue, FieldOrLab, Country) VALUES (\"{escape(row['Host'])}\", \"{escape(row['Group'])}\", \"{escape(row['HighTaxon'])}\", \"{escape(row['LowTaxon'])}\", \"{escape(row['Pathogen'])}\", \"0\", \"{escape(row['HostStageTested'])}\", \"NA\", \"Lab\", \"NA\")")
-
-with open('data/csv/nematode.csv', newline='', encoding='iso-8859-1') as csvfile:
-    reader = csv.DictReader(csvfile)
-    for row in reader:
-        c.execute(f"INSERT INTO Main (Host, ParasiteGroup, HighTaxon, LowTaxon, Pathogen, Infect, Stage, Tissue, FieldOrLab, Country) VALUES (\"{escape(row['Host'])}\", \"nematode\", \"{escape(row['NematodeOrder'])}\", \"{escape(row['NematodeFamily'])}\", \"{escape(row['Nematode'])}\", \"1\", \"{escape(row['StageInfected'])}\", \"{escape(row['TissueInfected'])}\", \"{escape(row['FieldOrLab'])}\", \"{escape(row['Country'])}\")")
-
-with open('data/csv/nvpassoc.csv', newline='', encoding='iso-8859-1') as csvfile:
-    reader = csv.DictReader(csvfile)
-    for row in reader:
[…]

### 2021-12-08 manuscript/webVersion/bin/form_values.py

@@ -1,55 +1,0 @@
-def query_list(field, db, table, args=()):
-    db.row_factory = lambda c, r : r[0]
-    cur = db.execute(f'SELECT {field} from {table}', args)
-    rows = [row for row in cur.fetchall() if row != 'no']
-    cur.close()
-    return sorted(list(set(rows)))
-
-def query_colnames(table, db, exclude=[], args=()):
-    db.row_factory = lambda c, r : r[0]
-    cur = db.execute(f'SELECT name FROM PRAGMA_TABLE_INFO(\'{table}\')')
-    cols = [col for col in cur.fetchall() if col not in exclude]
-    cur.close()
-    return sorted(list(set(cols)))
-
-def dropdowns(db, table='main'):
-    return {
-        'host_binomial'     : query_list('HostBinomial', db, 'Hosts'),
-        'host_order'        : query_list('HostOrder', db, 'Hosts'),
-        'host_family'       : query_list('HostFamily', db, 'Hosts'),
-        'host_genus'        : query_list('HostGenus', db, 'Hosts'),
-        'pathogen_name'     : query_list('Pathogen', db, 'Main'),
-        'pathogen_group'    : query_list('ParasiteGroup', db, 'Main'),
-        'status' : ['Both', 'Positive', 'Negative']
-    }
-
-def boxes():
-    return {
-        'host_fields' : ["Host Binomial (genus and species)"],
-        'parasite_fields' : ["Parasite Group (virus, bacteria, fungus, nematode, ect.)", "High Taxon (higher taxonomy)", "Low Taxon (lower taxonomy)", "Pathogen (genus and species)"],
-        'infection_fields' : ["Infect (infection status)", "Stage (host stage infected/ tested)", "Tissue (tissue infected/ tested)", "Field Or Lab (field or experimental observation)"],
-        'other_fields' : ["Country (country of host population)", "Host Order/ Taxon (higher host taxonomy)", "Host Family (lower host taxonomy)",
-            "Synonyms (alternate host scientific names)", "Habitat (host habitat)", "Food (host diet)",
-            "genYr (host generations per year)", "Common Name (host common name)", "Insect Status (pest or non-pest)"]
-    }
-
-columnMap = {
-    "hostbinomial(genusandspecies)"                         : "HostBinomial",
-    "parasitegroup(virus,bacteria,fungus,nematode,ect.)"    : "ParasiteGroup",
[…]

### 2021-12-08 manuscript/webVersion/bin/gunicorn.sh

@@ -1,3 +1,0 @@
-#!/bin/sh
-
-gunicorn -w 2 --threads 2 -b 0.0.0.0:5000 wsgi:app

### 2021-12-08 manuscript/webVersion/bin/query.py

@@ -1,36 +1,0 @@
-def makeQuery(data):
-    infectionStatus = data['InfectionStatus']
-    del(data['InfectionStatus'])
-
-    fieldlist = []
-    joinHosts = False
-    for field in [data[field][0] for field in data if '_fields_' in field]:
-        if field in ['HostBinomial', 'HostOrder', 'HostFamily', 'Synonyms', 'Habitat', 'Food', 'genYr', 'CommonName', 'InsectStatus']:
-            fieldlist.append(f'Hosts.{field}')
-            joinHosts = True
-        else:
-            fieldlist.append(f'Main.{field}')
-
-    fieldString = ', '.join(fieldlist)
-    if not fieldString:
-        fieldString = 'NULL'
-
-    queryString = f'SELECT {fieldString} FROM Main '
-    #if joinHosts:
-    queryString += f'INNER JOIN Hosts ON Main.Host = Hosts.HostBinomial '
-    wheres = ' AND '.join(['( '+' OR '.join([f'({field} = \'{value}\')' for value in data[field]])+ ' )' for field in data if '_fields_' not in field])
-    if wheres:
-        queryString += 'WHERE ' + wheres
-    if infectionStatus == 'Positive':
-        queryString += ' AND ( Infect = 1 )'
-    if infectionStatus == 'Negative':
-        queryString += ' AND ( Infect = 0 )'
-    return queryString
-
-def query(data, db):
-    query = makeQuery(data)
-    #print(query)
-    cur = db.execute(query,{})
-    records = cur.fetchall()
-    cur.close()
-    return [list(x) for x in records]

### 2021-12-08 manuscript/webVersion/bin/serve.py

@@ -1,65 +1,0 @@
-import sqlite3
-from flask import Flask, request, render_template, g, make_response
-
-import sys
-sys.path.append('./bin')
-
-import form_values
-import query
-
-app = Flask(__name__, root_path='./')
-
-def enquote(x):
-    if type(x) == str:
-        if ',' in x:
-            return f'"{x}"'
-        else:
-            return x
-    else:
-        return str(x)
-
-def make_dicts(cursor, row):
-    return dict((cursor.description[idx][0], value)
-                for idx, value in enumerate(row))
-
-def get_db(factory=None):
-    db = getattr(g, '_database', None)
-    if db is None:
-        db = g._database = sqlite3.connect('data/db.sqlite')
-        if factory:
-            db.row_factory = factory
-    return db
-
-@app.errorhandler(404)
-def page_not_found(e):
-    return render_template('404.html'), 404
-
-@app.errorhandler(500)
-def internal_error(e):
-    return render_template('500.html'), 500
-
-@app.route("/", methods=['GET'])
-def formPage():
-    return render_template('welcome.html', active='welcome'), 200
-    return '', 200
-
-@app.route("/download", methods=['GET', 'POST'])
-def queryPage():
-    if request.method == 'GET':
-        return render_template('form.html', dropdowns=form_values.dropdowns(get_db()), boxes=form_values.boxes(), active='download'), 200
-    if request.method == 'POST':
-        data = request.form.to_dict(flat=False)
-        for field in data:
-            if '_fields_' in field:
-                data[field] = [form_values.columnMap.get(data[field][0])]
-        #return(data)
-        results = query.query(data, get_db())
-        colnames = ','.join([data[field][0] for field in data if '_fields_' in field])
-        resp=make_response(colnames + '\n' + '\n'.join([','.join([enquote(x) for x in result]) for result in results]))
-        resp.headers.set('Content-Type', 'text/csv')
-        resp.headers.set("Content-Disposition", "attachment;filename=EDWIP_data.csv")
-        return(resp)
-    return render_template('500.html'), 500
-
-if __name__ == '__main__':
-    app.run(host='0.0.0.0',threaded=True,port=7778)

### 2021-12-08 manuscript/webVersion/static/js/bundle.js

@@ -1,79 +1,0 @@
-(function(){function r(e,n,t){function o(i,f){if(!n[i]){if(!e[i]){var c="function"==typeof require&&require;if(!f&&c)return c(i,!0);if(u)return u(i,!0);var a=new Error("Cannot find module '"+i+"'");throw a.code="MODULE_NOT_FOUND",a}var p=n[i]={exports:{}};e[i][0].call(p.exports,function(r){var n=e[i][1][r];return o(n||r)},p,p.exports,r,e,n,t)}return n[i].exports}for(var u="function"==typeof require&&require,i=0;i<t.length;i++)o(t[i]);return o}return r})()({"/Users/tomlin/RCN/parasites/parasites/static/js/index.js":[function(require,module,exports){
-'use strict';
-
-$(document).ready(function () {
-    $('#id_continent').on('hidden.bs.select', function (e, clickedIndex, newValue, oldValue) {
-        var continentSelected = $(this).val();
-        // hide options form country menu
-        var indexesToHide = [];
-        $('#id_country option').each(function (index) {
-            var continent = $(this).val().split('.')[0];
-            if (continentSelected && continentSelected.indexOf(continent) < 0) {
-                // if the continent for a country is not fund in the selected continents
-                // the country is hidden from selected.
-                indexesToHide.push(index);
-            }
-        });
[…]

### 2021-12-08 manuscript/webVersion/static/js/index.js

@@ -1,73 +1,0 @@
-$(document).ready(function() {
-    $('#id_continent').on('hidden.bs.select', function (e, clickedIndex, newValue, oldValue) {
-        var continentSelected = $(this).val();
-        // hide options form country menu
-        var indexesToHide = [];
-        $('#id_country option').each(function (index) {
-            var continent = $(this).val().split('.')[0];
-            if (continentSelected && continentSelected.indexOf(continent) < 0) {
-                // if the continent for a country is not fund in the selected continents
-                // the country is hidden from selected.
-                indexesToHide.push(index);
-            }
-        });
-        // go over widget options and hide as needed.
-        var $menu = $('#id_country').siblings('.dropdown-menu')[0];
-        $($menu).find('li').each(function(index) {
-           if (indexesToHide.indexOf(index) > -1) {
-               // this element needs to be hidden
-               $(this).addClass('hidden');
-               $(this).removeClass('selected');
-           } else {
-               // this element needs to be visible
-               $(this).removeClass('hidden');
-           }
-        });
-        // unselected previously selected values
-        var selectedCountries = $('#id_country').val(),
-            validSelectedCountries = [];
-        if (selectedCountries && selectedCountries.length) {
-            selectedCountries.forEach(function(item) {
-                var continent = item.split('.')[0];
-                if (continentSelected && continentSelected.indexOf(continent) > -1) {
-                    // country is in a selected continent so it is valid
-                    validSelectedCountries.push(item);
-                }
-            });
-            // only valid countries are set
-            $('#id_country').val(validSelectedCountries);
-            // check that the selected label is correct
-            var button = $('#id_country').siblings('button')[0],
-                selectWidgetLabel = $(button).find('.filter-option')[0],
-                currentSelection = $(selectWidgetLabel).text(),
-                labelStillValid;
-            if (validSelectedCountries) {
-                var numValidSelectedCountries = validSelectedCountries.length;
-                if (numValidSelectedCountries == 1) {
-                    // since there is only one valid country, let's display the country name
-                    $(button).attr('title', validSelectedCountries[0]);
-                    $(selectWidgetLabel).text(validSelectedCountries[0]);
-                } else {
-                    // display the number of items selected
-                    var labelMsg = numValidSelectedCountries + ' items selected';
-                    $(button).attr('title', labelMsg);
-                    $(selectWidgetLabel).text(labelMsg);
[…]

### 2021-12-08 manuscript/webVersion/static/libs/modernizr.js

@@ -1,1440 +1,0 @@
-/*!
- * modernizr v3.11.1
- * Build https://modernizr.com/download?-addtest-printshiv-setclasses-testprop-dontmin
- *
- * Copyright (c)
- *  Faruk Ates
- *  Paul Irish
- *  Alex Sexton
- *  Ryan Seddon
- *  Patrick Kettner
- *  Stu Cox
- *  Richard Herrera
- *  Veeck
-
- * MIT License
- */
-
-/*
- * Modernizr tests which native CSS3 and HTML5 features are available in the
- * current UA and makes the results available to you in two ways: as properties on
- * a global `Modernizr` object, and as classes on the `<html>` element. This
- * information allows you to progressively enhance your pages with a granular level
- * of control over the experience.
-*/
-
-;(function(scriptGlobalObject, window, document, undefined){
-
-  var tests = [];
-  
-
-  /**
-   * ModernizrProto is the constructor for Modernizr
-   *
-   * @class
-   * @access public
-   */
-  var ModernizrProto = {
-    // The current version, dummy
-    _version: '3.11.1',
-
-    // Any settings that don't work as separate modules
-    // can go in here as configuration.
-    _config: {
-      'classPrefix': '',
-      'enableClasses': true,
-      'enableJSClass': true,
-      'usePrefixes': true
-    },
-
-    // Queue of tests
-    _q: [],
-
-    // Stub these for people who are listening
-    on: function(test, cb) {
-      // I don't really think people should do this, but we can
-      // safe guard it a bit.
-      // -- NOTE:: this gets WAY overridden in src/addTest for actual async tests.
-      // This is in case people listen to synchronous tests. I would leave it out,
-      // but the code to *disallow* sync tests in the real version of this
-      // function is actually larger than this.
-      var self = this;
-      setTimeout(function() {
-        cb(self[test]);
-      }, 0);
-    },
-
-    addTest: function(name, fn, options) {
-      tests.push({name: name, fn: fn, options: options});
-    },
-
-    addAsyncTest: function(fn) {
-      tests.push({name: null, fn: fn});
-    }
-  };
-
-  
-
-  // Fake some of Object.create so we can force non test results to be non "own" properties.
-  var Modernizr = function() {};
-  Modernizr.prototype = ModernizrProto;
-
-  // Leak modernizr globally when you `require` it rather than force it here.
-  // Overwrite name so constructor name is nicer :D
-  Modernizr = new Modernizr();
-
-  
-
-  var classes = [];
-  
-
-  /**
[…]

### 2021-12-08 manuscript/webVersion/wsgi.py

@@ -1,1 +1,0 @@
-from bin.serve import app

### 2022-01-02 manuscript/manuscript.tex
almost ready

@@
 \author[a,b,*]{Tad A Dallas}
-\author[c]{Colin Carlson}
+\author[c,d]{Colin Carlson} % colin.carlson@georgetown.edu
+\author[e,f]{Patrick R Stephens} % prsteph@uga.edu
+\author[g, h, i]{Sadie J Ryan} % sjryan@ufl.edu
+\author[j]{David W Onstad} % dwonstad@iastate.edu
 
-% \author[c]{David Onstad}
-
-\affil[a]{Department of Biological Sciences, Louisiana State University, Baton Rouge, LA, USA}
-\affil[b]{Department of Biological Sciences, University of South Carolina, Columbia, SC, USA}
-
-\affil[c]{  }
-\affil[d]{  }
+\affil[a]{Department of Biological Sciences, Louisiana State University, Baton Rouge, LA, 70806}
+\affil[b]{Department of Biological Sciences, University of South Carolina, Columbia, SC, 29208}
+\affil[c]{Center for Global Health Science and Security, Georgetown University Medical Center, Washington, D.C. 20007}
+\affil[d]{Department of Microbiology and Immunology, Georgetown University Medical Center, Washington, D.C. 20007}
+\affil[e]{Center for the Ecology if Infectious Disease and Odum School of Ecology, University of Georgia, Athens, GA 30602}
+\affil[f]{Department of Integrative Biology, Oklahoma State University, Stillwater, OK 74078}
+\affil[g]{ Department of Geography, 3125 Turlington Hall, University of Florida, Gainesville, FL, 32611 USA.}
+\affil[h]{Emerging Pathogens Institute, P.O. Box 100009, 2055 Mowry Road, University of Florida, Gainesville, FL 32610 USA.}
+\affil[i]{School of Life Sciences, University of KwaZulu-Natal, Private Bag X54001, Durban, 4000, South Africa}
+\affil[j]{Corteva Agriscience, Johnston, IA, 50131}
 
@@
 %%% Target journal
-% Ecography (software note)
-%
+% Ecography (software note) -- maybe we shouldn't re-submit and pay $3000 to publish? 
+% Journal of Open Source Software 
+% Ecology (data paper)
 %
@@
 %%% Potential reviewers
-% 
-% 
+% Lewis Bartlett  lewis.Bartlett@uga.edu UGA
+% Philipp Boersch  pboesu@gmail.com   British Trust for Ornithology
+% Ben Longdon b.longdon2@exeter.ac.uk Uni of Exeter
+% Susan Braxton braxton@illinois.edu U Illinois (could be COI with Onstad though)
 % 
@@
 \begin{document}
+
 \maketitle
+
+\clearpage 
 \setstretch{1}
@@
 
-\noindent \textbf{Acknowledgements}: We sincerely thank all individuals associated with the collection and curation of the \textit{Ecological Database of the World's Insect Pathogens} data, specifically David Onstad at the US Department of Agriculture.
+\noindent \textbf{Acknowledgements}:  This work was supported by funding to the Viral Emergence Research Initiative (VERENA) consortium including NSF BII 2021909.
+For more information on the Verena Consortium, see \texttt{https://viralemergence.org}.
+Earlier development of these data was supported by the Macroecology of Infectious Disease Research Coordination Network (NSF/NIH/USDA DEB 131223).
 \\
@@
 
-\noindent \textbf{Keywords}: insect pathogens, crop pest, experimental infection,   \\
+\noindent \textbf{Keywords}: insect pathogens, crop pest, experimental infection, nematode, virus, protozoa, bacteria  \\
 
@@
 \subsection*{Abstract}
+Curated databases of species interactions are instrumental to exploring and understanding the spatial distribution of species and their biotic interactions.
+In the process of conducting such projects, data development and curation efforts may give rise to a data product with utility beyond the scope of the original work, but which becomes inaccessible over time.
+Data describing insect host-pathogen interactions are fairly rare, and should thus be preserved and curated with appropriate metadata.
+Here, we introduce the \texttt{insectDisease} $R$ package, a mechanism for curating, updating, and distributing data from the \textit{Ecological Database of the World's Insect Pathogens}, a database of insect host-pathogen associations, including attempted inoculations and infection outcomes for insect hosts and pathogens (bacteria, fungi, nematodes, protozoans, and viruses).
+This dataset has been utilized for several projects since its inception, but without a well-defined, curated and permanent repository, its existence and access have been limited to word-of-mouth connections.
+The current effort presented here aims to provide a means to preserve, augment, and disseminate the database in a documented and versioned format.
+This project is an example of the type of effort that will be necessary to maintain valuable databases after the original funding disappears.
 
@@
 There are a number of data sources documenting host-pathogen associations, especially for pathogens of mammals \citep{gibb2021, patrick2017}, birds \citep{bensch2009}, and fish \citep{fishpest}.
-These have been incredibly important to our understanding of what determines pathogen host range, pathogen species richness across a set of hosts, and overall host-pathogen network structure (e.g, \citep{dallas2018, carlson2020}).
+Recent work from the Verena Consortium has developed a dynamically updated host-virus association database for all vertebrate hosts (\texttt{VIRION}) \citep{carlson2021}, representing the largest collection of host-virus association data to date.
+These resources have been fundamental to our understanding of what determines pathogen host range, pathogen species richness across a set of hosts, and overall host-pathogen network structure (e.g, \citet{dallas2018, carlson2020}).
 But while some host groups are well-studied, there are taxonomic gaps in our understanding of host-pathogen associations.
-For instance, insect host-pathogen relationships have considerably less open-source data available, despite their inherent importance to agricultural crops and vector-borne disease, in addition to the numerical dominance of insect species over other taxa \citep{stork2015}.
+Insect host-pathogen relationships have considerably less open-source data available, despite their inherent importance to scientific studies and assessments of impacts to agricultural crops and spread of vector-borne disease, in addition to the sheer numerical dominance of insect species over other taxa \citep{stork2015}.
 This is a clear knowledge gap.
 
-% the file drawer issue and
 \paragraph*{}
-Additionally, all of the previous data sources listed before have dedicated researchers and government agencies responsible for data deposition and curation in openly accessible formats.
+Many of the existing species interaction databases have dedicated researchers, resources, and infrastructure to enable data deposition and curation in openly accessible formats.
 However, some data have not been as lucky, at no fault of the original data curators.
@@
 \paragraph*{}
-The EDWIP data consist of experimental infections and field observations of the interactions between insect hosts and a number of nematode, viral, protozoan, fungal, and bacterial pathogens \citep{braxton2003}.
-In the EDWIP data, a number of pathogen infection types were excluded, including vector-borne pathogens which do not infect the insect host.
-An interesting component of EDWIP is the existence of negative associations -- attempts to inoculate a host with a given pathogen that failed to infect -- for some host groups (Figure \ref{fig:bubble}).
+The EDWIP data consist of experimental infections and field observations of the interactions between insect hosts and a number of bacterial, fungal, nematode, protozoan, and viral pathogens \citep{braxton2003}.
+One particularly unique component of EDWIP is the existence of negative associations -- attempts to inoculate a host with a given pathogen that failed to infect -- for some host groups (Figure \ref{fig:bubble}).
 Failed infections represent \textit{true} absences or incompatibilities between a given host and pathogen.
-These data are incredibly useful to pathogen host range and host-pathogen interaction modeling, but we rarely have data on these known non-interactions.
-Initially created in 1996, the data have been dynamically updated over time.
-However, it is unclear how long this curation continued, and how many different versions of the data may be in existence without proper versioning.
-To this point, the numbers of insect-host pathogen interactions differs slightly from previously published versions of the data \citep{braxton2003}, but these changes are relatively minor.
+These data are incredibly useful to pathogen host range estimation and host-pathogen interaction modeling, but we rarely have data on these known non-interactions.
+
+\paragraph*{}
+Initially created in 1992, the data have been updated prior to 2000, but no clear semantic versioning was used.
+As such, it is unclear how long or how frequently this updating and curation continued, and thus, how many different versions of the data may be in existence presently.
+The database we present here, as the backbone of this R package, represents the most up-to-date version that we know of, though this may differ slightly from previous descriptions of the data \citep{braxton2003}.
+Generally, we have attempted to preserve all of the original data in the original format.
 
 \subsection*{Solution statement}
-
 To preserve these data in a format that is well-documented, openly accessible, versioned, and flexible for continued development, we created the \texttt{insectDisease} R package.
-In doing so, we implicitly adhere to the FAIR (Findable, Accessible, Interoperable, Resuable) guidelines for managing data.
+In doing so, we implicitly adhere to the FAIR (Findable, Accessible, Interoperable, Resuable) guidelines for managing data \citep{wilkinson2016}.
 By hosting the data openly on GitHub, and versioning releases of the data with a permanent identifier (DOI), we ensure the longevity and versioned curation of this data resource.
-Finally, the incorporation of taxonomic data through \texttt{taxize} \citep{chamberlain2013} makes sure that host and pathogen taxonomic names are updated.
+Finally, the incorporation of taxonomic data through \texttt{taxize} \citep{chamberlain2013} ensures that host and pathogen taxonomic names are updated periodically to accommodate for dynamic data or changing taxonomies.
 
@@
 
-% Discuss more of the in-depth view of the R package as a means for data delivery, and what the overall workflow looks like.
-Discuss how it's open and all part of the Verena ecosystem (or whatever).
-Flex on data size and point to figures displaying the positive and negative associations and such.
-
 \paragraph*{Package structure}
-
 Data products are broken down by pathogen group; nematodes (\texttt{data(nematode)}), viruses (\texttt{data(viruses)}), and non-viral pathogens, which include protozoan, fungi, and bacteria (\texttt{data(nvpassoc)}).
-Data on negative associations is stored collectively instead of being delineated by pathogen group (\texttt{data(negative)}), but information on pathogen group is provided within each of these files, allowing for sorting of negative interactions based on the initial pathogen groupings.
+Data on negative associations is stored collectively instead of being delineated by pathogen group (\texttt{data(negative)}), but information on pathogen group is provided within each of these files, allowing for sorting of negative interactions based on the initial pathogen groupings (Table \ref{tab:data}).
+This data structure is inherited from the original structure of the EDWIP data files, and code to process and join these different data files is provided in the $R$ package vignette.
 
 \paragraph*{}
-Each of the pathogen groups slightly differs in the available data on experimental infections.
-For instance, nematode infections contain information on soil type and associated bacteria, virus infection data has information on viral dose, and non-viral pathogens (protozoans, fungi, and bacteria) have information on intermediate host species.
+Each of the pathogen groups differs slightly in the available ancillary data on experimental infections.
+For instance, nematode infections contain information on soil type and associated bacteria, virus infection data have information on viral dose, and non-viral pathogens (protozoans, fungi, and bacteria) have information on intermediate host species.
+We recommend the user explore these data and associated metadata from within $R$, as the metadata and data are neatly in the same place.
 
@@
 Data are also available on the insect host species themselves (e.g., \texttt{data(hosts)}).
-These data contain some information on Canadian province where the host is found (\texttt{ProvinceI} column), what it eats (\texttt{Food} column), and what type of habitat it is found in (\texttt{Habitat} column).
+These data contain some information on the Canadian province where the host is found (\texttt{ProvinceI} column), what it eats (\texttt{Food} column), and what type of habitat it is found in (\texttt{Habitat} column).
 Additionally, a column on host insect pest status is present, offering the opportunity to explore study effort and pathogen specificity dependent on the pest status of the insect host.
+
+\paragraph*{FAIR data}
+The FAIR principles represent guidelines for making data more persistent, findable, and well-documented.
+Structuring the data as an R package ensures that metadata and data are packaged together, where R manual files contain column names and data descriptions for each data product (\textit{Findable}).
+All code to take data from the raw data (\texttt{data-raw} folder) to the end product .RData and .csv files is contained in the versioned R data package, and integration with Zenodo provides a DOI for each release (\textit{Accessible}).
+Apart from providing data in these multiple formats, user access is aided by structuring the data as a package in a very popular computing language among biologists (and other folks too) and providing all code for data processing and serving in an open and public-facing repository (\textit{Interoperable}).
+Having all code and data in a streamlined, open, and versioned format, serving the data through an interactive web portal, and publishing this software note collectively serve to promote the use of this data resource (\textit{Reusable}).
 
@@
 
-Differences in features across the data on different pathogen types (e.g., \texttt{?nematodes} relative to \texttt{?viruses}) make combining these data difficult without a loss of information.
-However, there is clear utility in having all of the data in a single standardized form.
-As such, we have documented each data resource using R package documentation, allowing the metadata of each data product to be examined directly from R using the \texttt{help()} function or the question mark notation (e.g., \texttt{?viruses}).
+Differences in features across the data on different pathogen types (e.g., \texttt{?nematodes} relative to \texttt{?viruses}) make combining these data non-straightforward, without a degree of loss of information.
+We provide some example code in the package \texttt{vignette} on how to go about combining or linking the data across types, with the caveats of information loss, and have standardized some key column names across the different data products.
+Further, we have documented each data resource using $R$ package documentation, allowing the metadata of each data product to be examined directly from R using the \texttt{help()} function or the question mark notation (e.g., \texttt{?viruses}).
 
@@
 We attempted to maintain as much of the original data structure from the raw data files provided by David Onstad, principal maintainer of the EDWIP data resource \citep{onstad1997}.
-This includes files such as \texttt{new\_assoc}, as this was likely a test file containing pathogen species such as "wormy thing".
-We provide comments in the documentation about which data sources are perhaps more trustworthy.
-We clean and augment this existing data source programmatically, with much of this code in the \texttt{insectDisease} package \texttt{vignette}.
+This includes files such as \texttt{new\_assoc}, as this was likely a test file containing pathogen species such as ``wormy thing'', and \texttt{newnema}, a dataset identical to \texttt{nematode}.
+We document these idiosyncrasies in the metadata for each data product, providing a clear overview of the state of each data subproduct.
 
 \paragraph*{}
-The first, and perhaps most important augmentation, is the resolution of host and pathogen taxonomic information.
-We achieved this by using the R package \texttt{taxize}, specifically the NCBI taxonomic backbone \citep{chamberlain2013}.
-Cached versions of host and pathogen taxonomic information are provided (\texttt{data(hostTaxonomy)} and \texttt{data(pathTaxonomy)}), and the $R$ code to generate these taxonomic backbones and clean the data are provided in the package vignette.
+The first, and perhaps most important, novel augmentation, is the resolution of host and pathogen taxonomic information.
+We achieved this by using the R package \texttt{taxize}, specifically the NCBI taxonomic backbone \citep{chamberlain2013}, making the data interoperable with existing data efforts by the Verena Consortium (e.g., VIRION; \citet{carlson2021}).
+Cached versions of host and pathogen taxonomic information are provided (\texttt{data(hostTaxonomy)} and \\  \texttt{data(pathTaxonomy)}), and the $R$ code to generate these taxonomic backbones and clean the data are provided in the package vignette.
 This taxonomic backbone serves to both standardize host and pathogen nomenclature, while also correcting any taxonomic changes that have occurred in the past couple decades.
 This includes the consideration of microsporidian parasites as fungi, not protozoans, a change affecting a large set of records in the EDWIP data.
-Second, we provide a vignette which transparently shows the steps to clean and augment the data resource, as well as reproduce figures from this manuscript.
-Further, the vignette includes code which generates host-pathogen associations for all the data combined into a single data resource, simplifying the data down to a small number of important columns.
-Finally, processed we opt to store raw data in the \texttt{csv} folder, which contains all raw data files in .csv format.
-These data are also provided as .rda files in the \texttt{data} folder.
+All of the data within the \texttt{data} and \texttt{csv} folders have already gone through these data cleaning steps.
+However, these data may be dynamic, such that some form of continuous integration or updating of the host and pathogen taxonomy may be necessary.
+As such, we provide a vignette which transparently shows the steps to clean and augment the data resource, as well as reproduce figures from this manuscript.
+Finally, we opt to store processed data in the \texttt{csv} folder, which contains all data files in  \texttt{.csv} format.
 This allows non-$R$ users to access the csv-formatted data easily, and ensures long-term stability of the data, as \texttt{csv} is a stable text file format.
+These data are also provided as \texttt{.rda} files in the \texttt{data} folder.
+
+\paragraph*{}
+Maintaining the data dynamically as described above allows users to access the data programmatically or as versioned flatfiles (i.e., \texttt{.csv} files).
+However, for users who do not wish to download the entire data resource, and simply want to quickly query a static version of the database, there is also a standalone web user interface (\texttt{https://edwip.ecology.uga.edu/}) that allows users to easily subset and explore the data.
+The web interface serves arguably the most important subset of the overall data (data files \texttt{nematode}, \texttt{viruses},\texttt{nvpassoc}, \texttt{negative}, and \texttt{hosts}).
+This interface allows users to quickly query based on host or parasite taxonomy as a dropdown list.
+This is perhaps more useful as a teaching tool or for initial exploration of the data, while the programmatic interface and dynamic data may be more useful for more rigorous analysis.
+This version of the EDWIP data will also only be deployed with a single static copy of the data, such that users wanting to benefit from versioned and dynamic data will need to access the data through the GitHub repository.
+Future efforts to integrate the web interface and the existing dynamic data structure will be explored, but this is not currently integrated.
 
@@
 We explore this in the EDWIP data by measuring the number of known positive associations of each of the pathogen groups for each insect host species, visualizing the relationship between the number of pathogens per insect host as a correlation matrix (Figure \ref{fig:corPlot}).
-We find very little evidence that pathogen groups have positive covariance, which would be expected if host species traits drove infection process across pathogen groups in the same manner.
-The failure to detect positive relationships, and indeed some negative relationships appearing, could be a signal of the targeted nature of data collection, as many insect host species were selected to study due to their potential as a crop pest, and many pathogens were selected to study based on their potential use as biocontrol or perhaps for their ease of culture.
+We find very little evidence that pathogen groups have positive covariance, which would be expected if host species traits or trait-based sampling biases drove infection process across pathogen groups in the same manner.
+The failure to detect strong positive relationships, and indeed some negative relationships appearing, could be a signal of the targeted nature of data collection, as many insect host species were selected to study due to their potential as a crop pest, and many pathogens were selected to study based on their potential use as biocontrol or perhaps for their ease of culture.
 
 \paragraph*{}
-This potential sampling bias among insect host species would be clear if there were a positive relationship between the number of positive interactions to the number of negative interactions for a host species, as it would indicate that insect host species with lots of known interactions also tended to appear in many studies and have some negative interactions as well.
-We find evidence for a significantly negative relationship based on a Spearman's rank correlation ($\rho$ = -0.1, p $<$ 0.0001).
+This potential sampling bias among insect host species would be evident if there were a positive relationship between the number of positive interactions and the number of negative interactions for a host species, as it would indicate that host species with lots of known interactions also tended to appear in many studies and have some negative interactions as well.
+We find evidence for a significantly negative relationship based on a Spearman's rank correlation ($\rho$ = -0.1, p $<$ 0.0001), indicating no discernible influence of this relationship.
+This does not imply that there is no sampling bias in the insect host species researchers opt to study, but that such bias was not so strong as to be clearly detected.
 
@@
 
-Ecological data -- while growing in availability, size, and stability -- are still a resource that should not be allowed to fade out of existence.
-The EDWIP data provided to the authors were in a proprietary format ('Claris FileMaker Pro 5') that was already over 10 major versions behind.
+While ecological data are growing in availability, size, accessibility, and stability, there are still data resources that are aging in place, and should not be allowed to fade out of existence.
+The EDWIP data provided to the authors were in a proprietary format (`Claris FileMaker Pro 5') that was already over 10 major versions behind.
 With limited inter-version operability (e.g., \texttt{.fmp5} files cannot be opened in more recent versions of the software, or require multiple conversion steps), these data seemed as if headed towards obsolescence.
@@
 \subsection*{Data accessibility}
-
-The \texttt{insectDisease} R package is currently available on GitHub https://github.com/viralemergence/insectDisease, with '.csv' files in the csv directory for long-term data stability.
-
-Finally, the data are periodically archived at major version changes via Zenodo ??????????????????
+The \texttt{insectDisease} R package is currently available on GitHub \\ (github.com/viralemergence/insectDisease), with `.csv' files in the csv directory for long-term data stability.
+GitHub releases of the data ensure versioning is maintained and all versions are accessible.
+At the time of this writing, the current version is 1.1.0 (available at \\ \texttt{https://github.com/viralemergence/insectDisease/releases/tag/1.1.0}).
 
@@
 %----------------
+
+\section*{Tables}
+
+\begin{table}[ht]
+\centering
+\caption{Files associated with the EDWIP data resource.
+Metadata is stored in $R$ package documentation, allowing the data and metadata to be intrinsically linked.
+For instance, users can use the help functionality from within $R$ to see more information on data columns and unit (e.g., \texttt{?nematode}).
+}
+\label{tab:data}
+\begin{tabular}{cccp{8cm}}
+  \hline
+  filename & rows & columns & description \\
+  \hline
+   \texttt{assocref} & 11005  & 16 & references for some host-pathogen associations  \\
+   \texttt{citation} & 1966  & 7  & references but no host-pathogen association information  \\
+   \texttt{hosts} & 4392  & 21  & insect host trait data\\   
+   \texttt{hostTaxonomy} & 4489  & 7  & host taxonomic data updated with the \texttt{getNCBI()} function \\
+   \texttt{negative} & 529  & 21  & information on negative host-pathogen associations  \\
+   \texttt{nemaref} & 338  & 5  & references from nematode pathogens  \\
+   \texttt{nematode} & 234  & 24  & host-nematode interaction data \\   
+   \texttt{new\_asso} & 19  & 25  & likely a training document (perhaps do not use)  \\ 
+   \texttt{noassref} & 569  & 16  &  references for some host-pathogen associations  \\
+   \texttt{nvpassoc} & 7164  & 23  & non-viral pathogen infection data  \\   
+   \texttt{pathogen} & 2041  & 9  &  pathogen trait data  \\   
+   \texttt{pathTaxonomy} & 2282  & 7  & pathogen taxonomic data updated with the \texttt{getNCBI()} function  \\
+   \texttt{viraref} & 2124  & 16  & references from viral infections  \\   
+   \texttt{viruses} & 1659  & 25  & host-viral interaction data  \\      
+  \hline
+\end{tabular}
+\end{table}
+
+\clearpage
+
 \section*{Figures}
@@
     \includegraphics[width=0.85\textwidth]{Figures/bubble.pdf}
-    \caption{Bubble plot, where points are proportional to the total number of negative (on the left) and positive (on the right) host-pathogen interactions for each pathogen group ($y$-axis).
+    \caption{The number of known non-interactions (\textit{negative} left panel) and known interactions (\textit{positive} right panel) for the set of bacterial, fungal, nematode, protozoan, and viral pathogens ($y$-axis).
+Bubble size is proportional to the total number of interactions associated with that pathogen group and interaction type (i.e., \textit{negative} or \textit{positive}).
 Numeric columns correspond to the number of unique host species, pathogen species, and interactions for each pathogen group.
@@
     \caption{Correlations between each pathogen group in terms of pathogen richness of insect host species, where color corresponds to Spearman's rank correlation values (provided in the lower diagonal matrix).
-Fungal and protozoan pathogens seemed to negatively covary, as did viruses and protozoans.
+Fungal and protozoan pathogens were negatively related, as were viruses and protozoans.
 Understanding to what extent this is driven by sampling effects or insect host ecology is an outstanding research question that these could be used to begin addressing.}

### 2022-01-02 data-raw/process.R

@@ -194,6 +194,106 @@ viruses  <- fixDates(viruses)
 
 
 
+
+
+
+#' check for weird \v characters
+#' 
+#' @param data 
+#'
+#' @return updated data.frame
+
+fixV <- function(data){
+
+  inds <- apply(data, 2, function(x){any(grep('\v', x))})
+  if(any(inds==TRUE)){
+    inds <- which(inds)
+    for(i in 1:length(inds)){
+      # host species had some \v characters too caused by unicode garbo and bad data formatting
+      if(names(inds)[i] %in% c('HostSpecies', 'PathogenSpecies', 'Reference', 'ChangeSpeciesTo', 'LogMaxDose', 'NemaOrder', 'Citation', 'AdditionalNotes', 'Virus', 'Country') ){
+        data[,inds[i]] <- gsub('\v', '', data[,inds[i]])    
+      }else{
+      # but the focus is on providing separation for multiple entry columns like HostStageInfected
+       data[,inds[i]] <- gsub('\v', ';', data[,inds[i]])
+      }
+    }
+  }
+  return(data)
+}
+
+
+
+
+assocref <- fixV(assocref)
+citation <- fixV(citation)
+hosts    <- fixV(hosts)
+negative <- fixV(negative)
+nemaref  <- fixV(nemaref)
+nematode <- fixV(nematode)
+new_asso <- fixV(new_asso)
+newnema  <- fixV(newnema)
+noassref <- fixV(noassref)
+nvpassoc <- fixV(nvpassoc)
+pathogen <- fixV(pathogen)
+viraref  <- fixV(viraref)
+viruses  <- fixV(viruses)
+
+
+
+
+
+
+#' insert NA where data were just empty
+#' 
+#' @param data 
+#'
+#' @return updated data.frame
+
+fixNA <- function(data){
+  data <- as.data.frame(
+    lapply(data, function(x){
+      if(any(nchar(x)==0, na.rm=TRUE)){
+        x[which(nchar(x)==0)] <- NA
+      }
+      return(x)
+    })
+  )
+  return(data)
+}
+
+
+
+
+assocref <- fixNA(assocref)
+citation <- fixNA(citation)
+hosts    <- fixNA(hosts)
+negative <- fixNA(negative)
+nemaref  <- fixNA(nemaref)
+nematode <- fixNA(nematode)
+new_asso <- fixNA(new_asso)
[…]

### 2022-01-02 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+test_check("insectDisease")
+
+

### 2022-01-02 tests/testthat/test-negative.R

@@ -1,0 +1,14 @@
+context("negative")
+
+test_that("", {
+  skip_on_cran()
+  data(negative)
+	expect_is(negative, "data.frame")
+	expect_true(class(negative$CreationDate) == 'Date')
+  expect_true(class(negative$ModificationDate) == 'Date')
+  expect_true(all(negative$ModificationDate < "2021-01-01"))
+})
+
+
+
+

### 2022-01-02 tests/testthat/test-nematodes.R

@@ -1,0 +1,14 @@
+context("nematode")
+
+test_that("wormy time", {
+  skip_on_cran()
+  data(nematode)
+	expect_is(nematode, "data.frame")
+	expect_true(class(nematode$CreationDate) == 'Date')
+  expect_true(class(nematode$ModificationDate) == 'Date')
+  expect_true(all(nematode$ModificationDate < "2021-01-01"))
+})
+
+
+
+

### 2022-01-02 tests/testthat/test-nvpassoc.R

@@ -1,0 +1,11 @@
+context("nvpassoc")
+
+test_that("", {
+  skip_on_cran()
+  data(nvpassoc)
+	expect_is(nvpassoc, "data.frame")
+})
+
+
+
+

### 2022-01-02 tests/testthat/test-viruses.R

@@ -1,0 +1,13 @@
+context("viruses")
+
+test_that("", {
+  skip_on_cran()
+  data(viruses)
+	expect_true(class(viruses$CreationDate) == 'Date')
+  expect_true(class(viruses$ModificationDate) == 'Date')
+  expect_true(all(viruses$ModificationDate < "2021-01-01"))
+})
+
+
+
+

### 2022-01-05 manuscript/coverLetter/Letter.tex
updated and prepped for release and zenodo-ing

@@
 \usepackage{amssymb}
-\usepackage{setspace}
 \usepackage{amsthm}
@@
 \begin{flushleft}
-	\begin{minipage}[c]{4cm}
+	\begin{minipage}[c]{6cm}
 		\begin{flushleft}
-			\includegraphics*[width=4cm]{lsu.png}%
+			\includegraphics*[width=6cm]{usc.jpg}%
 		\end{flushleft}
@@
 			Department of Biological Sciences \\
-			Louisiana State University \\
-			Baton Rouge, LA 70802 \\
-			phone: +1 225 239 0410  \\
-			\email{tad.a.dallas@gmail.com}
+			University of South Carolina\\
+			Columbia, SC 29208 \\
+			\email{tdallas@mailbox.sc.edu}
 		}
@@
 
-Please find the enclosed manuscript ``A latitudinal signal in the relationship between species geographic range size and climatic niche area'', which we submit for consideration as a research article.
+Please find the enclosed manuscript ``\texttt{insectDisease}: programmatic access to the \textit{Ecological Database of the World's Insect Pathogens}'', which we re-submit for consideration as a research article.
 \\
 
-Species with wider climatic niche breadths are expected to be more widely distributed geographically.
-But while there appears to be strong evidence for this relationship, an understanding of when deviations would be expected as a result of geographic effects or species traits is an open question.
-Here, we use data on over 3000 species of mammal, bird, and tree, relating the model residuals from the relationship between geographic range size and climatic niche area to species traits and geographic position, and compare these residuals to what would be expected under a null model.
+Curated databases of species interactions are instrumental to exploring and understanding the spatial distribution of species and their biotic interactions.
+In the process of conducting such projects, data development and curation efforts may give rise to a data product with utility beyond the scope of the original work, but which becomes inaccessible over time.
+We present one such example, the "Ecological Database of World's Insect Pathogens".
+The R data package and manuscript text have benefitted from the comments of two reviewers and editor Dr.
+Michael Borregaard.
+These comments centered on the package structure and the presentation of the data.
+Specifically, the data required processing, but we were unaware of the current standard practice of having a \texttt{data-raw} folder and a processing script.
+We have incorporated this, and made more data checks -- including the incorporation of unit tests in the pacakge -- to handle odd date issues, remove bad unicode characters, and provide some signal of potential error as these data will be augmented in the future with more records.
+The comments from all reviewers seemed to indicate that substantial revision was necessary, and we believe we have made substantial modifications to both manuscript text and R package structure.
+We thank the reviewers and the editor for their critiques.
 \\
 
-We find evidence for an effect of species latitudinal centre on model residuals, suggesting the existence of a consistent spatial signal in deviation from range size -- climatic niche area scaling relationships across a diverse set of species.
-Some species traits were also important (e.g., body size for mammals), but the distribution of model residuals comparing empirical to null relationship did not differ appreciably.
-Together, our analyses suggest that geographic range size -- climatic niche area relationships are more a function of the landscape and a spatially autocorrelated environment than driven by species traits.
-\\
-
-We believe this article will be of interest to a broad range of ecologists and biogeographers, as it addresses a longstanding question in ecology concerning the link between species niche and distribution.
+We believe this article will be of interest to a broad range of ecologists interested in pest management, disease, species interactions, and more.
 In advance, we appreciate your consideration and assistance.

### 2022-01-05 manuscript/manuscript.tex
zenodo info added to manuscript

@@
 Structuring the data as an R package ensures that metadata and data are packaged together, where R manual files contain column names and data descriptions for each data product (\textit{Findable}).
-All code to take data from the raw data (\texttt{data-raw} folder) to the end product .RData and .csv files is contained in the versioned R data package, and integration with Zenodo provides a DOI for each release (\textit{Accessible}).
+All code to take data from the raw data (\texttt{data-raw} folder) to the end product .RData and .csv files is contained in the versioned R data package, and integration with Zenodo (\texttt{https://doi.org/10.5281/zenodo.5821896}) provides a DOI for each release (\textit{Accessible}).
 Apart from providing data in these multiple formats, user access is aided by structuring the data as a package in a very popular computing language among biologists (and other folks too) and providing all code for data processing and serving in an open and public-facing repository (\textit{Interoperable}).
@@
 GitHub releases of the data ensure versioning is maintained and all versions are accessible.
-At the time of this writing, the current version is 1.1.0 (available at \\ \texttt{https://github.com/viralemergence/insectDisease/releases/tag/1.1.0}).
+At the time of this writing, the current version is 1.2.0 (available at \\ \texttt{https://github.com/viralemergence/insectDisease/releases/tag/1.2.0}).
+Releases are given a DOI through integration with Zenodo (\texttt{https://doi.org/10.5281/zenodo.5821896}).
 

### 2022-04-18 manuscript/manuscript.tex
second round revisions at ecography

@@
 All code to take data from the raw data (\texttt{data-raw} folder) to the end product .RData and .csv files is contained in the versioned R data package, and integration with Zenodo (\texttt{https://doi.org/10.5281/zenodo.5821896}) provides a DOI for each release (\textit{Accessible}).
+Metadata are available in redundant forms, both from within the R package as \texttt{man} files, and in the project README file such that installation of the package (or navigation into the \texttt{man} folder) is not necessary.
 Apart from providing data in these multiple formats, user access is aided by structuring the data as a package in a very popular computing language among biologists (and other folks too) and providing all code for data processing and serving in an open and public-facing repository (\textit{Interoperable}).
@@
 Further, we have documented each data resource using $R$ package documentation, allowing the metadata of each data product to be examined directly from R using the \texttt{help()} function or the question mark notation (e.g., \texttt{?viruses}).
+These same metadata are also provided in the README file in the top-level of the GitHub repository.
 

### 2022-04-18 R/pathTaxonomy.R

@@ -8,7 +8,7 @@
 #'   \item PathFamily pathogen family
 #'   \item PathOrder pathogen order
 #'   \item PathClass  pathogen class
-#'   \item PathKingdom  pathogen class
+#'   \item PathKingdom  pathogen kingdom
 #'  }
 #'
 #' @docType data

### 2022-04-18 vignettes/dataManip.R

@@ -1,8 +1,0 @@
-## ---- results='hide'----------------------------------------------------------
-
-library(insectDisease)
-library(plyr)
-library(dplyr)
-library(corrplot)
-knitr::opts_chunk$set(fig.width=6, fig.height=6) 
-

### 2022-06-13 manuscript/coverLetter/Letter.tex
second round revisions at ecography
updates to readme and manuscript edits

@@
 
-{\noindent Dear \textit{Ecology Letters} editorial board,} \vskip4pt
+{\noindent Dear \textit{Ecography} editorial board,} \vskip4pt
 
@@
 We thank the reviewers and the editor for their critiques.
+\\
+
+I (the lead author, as I do not mean to speak for the co-authors on this point) am honestly curious in continuing this conversation, as there is a lot of nuance here.
+For instance, I thought having the documentation as plain text (.Rd) files in the `man` folder would still qualify as FAIR since it is readable without having to install the package.
+This is a more minor point to my larger curiosity about the shifting landscape of package development, and how the tools for making packages seem to enforce a rigidity in dev and toolsets (e.g., pkgdown, etc.) outside of `devtools` (and even `devtools` is less a single package and more of a collection of packages now with a set of 19+ dependencies).
+What will this mean to future developers?
+For the future of CRAN or centralized repositories?
 \\

### 2022-06-13 R/citation.R

@@ -10,9 +10,9 @@
 #'   \item Reference. Actual citation
 #'   \item CreationDate. Record creation date
 #'   \item ModificationDate. Record modification date
-#'   \item ReadBy. 
-#'   \item GetIt. 
-#'   \item nvpCount. 
+#'   \item ReadBy. Comments about the reading and identity of reader 
+#'   \item GetIt. Notes on article acquisition    
+#'   \item nvpCount. Number of pathogens reported in the citation.
 #'  }
 #'
 #' @docType data

### 2022-08-08 R/assocref.R
updates based on reviewer comments

@@ -26,7 +26,7 @@
 #' @keywords datasets
 #' @name assocref
 #' @usage data(assocref)
-#' @format A data.frame with 11005 observations of 5 variables
+#' @format A data.frame with 11005 observations 
 
 
 

### 2022-08-08 R/citation.R

@@ -19,7 +19,7 @@
 #' @keywords datasets
 #' @name citation
 #' @usage data(citation)
-#' @format A data.frame with 1966 observations of 7 variables
+#' @format A data.frame with 1966 observations
 
 
 

### 2022-08-08 R/hosts.R

@@ -10,8 +10,6 @@
 #'   \item DateEntered. Date of initial data entry
 #'   \item Habitat. Habitat of host
 #'   \item HostSpecies. Host species
-#'   \item HostOrder. Host order
-#'   \item HostFamily. Host family
 #'   \item Synonyms. Other names for the host species
 #'   \item Food. What does the host eat?
 #'   \item genYr. Number of generations per year. Can be <1, =1, >1, or some combination of these.
@@ -23,7 +21,7 @@
 #'   \item InCanada. Citations for presence/absence of host in Canada. Numeric indices can be
 #'     related to the citations in the `citations.rda` data file. `Y` and `N` 
 #'     relate to presence and absence, respectively.
-#'   \item ChangeSpeciesTo. 
+#'   \item ChangeSpeciesTo. Taxonomic verification column
 #'   \item CommonNameOther. Other common names?
 #'   \item Complete. Is this record complete?
 #'   \item AdditionalReferences. Additional reference indices. 
@@ -38,7 +36,7 @@
 #' @keywords datasets
 #' @name hosts
 #' @usage data(hosts)
-#' @format A data.frame with 4392 observations of 18 variables
+#' @format A data.frame with 4392 observations 
 
 
 

### 2022-08-08 R/negative.R

@@ -34,6 +34,6 @@
 #' @keywords datasets
 #' @name negative
 #' @usage data(negative)
-#' @format A data.frame with 529 observations of 12 variables
+#' @format A data.frame with 529 observations 
 
 "negative"

### 2022-08-08 R/nemaref.R

@@ -27,7 +27,7 @@
 #' @keywords datasets
 #' @name nemaref
 #' @usage data(nemaref)
-#' @format A data.frame with 338 observations of 5 variables
+#' @format A data.frame with 338 observations
 
 
 

### 2022-08-08 R/nematode.R

@@ -4,7 +4,7 @@
 #' 
 #' 
 #' \itemize{
-#'   \item ERNnem
+#'   \item ERNnem.  EDWIP record identifier
 #'   \item HostSpecies. Host species
 #'   \item PathogenSpecies. Nematode parasite species
 #'   \item PathogenStrain. Nematode parasite strain
@@ -34,6 +34,6 @@
 #' @keywords datasets
 #' @name nematode
 #' @usage data(nematode)
-#' @format A data.frame with 234 observations of 15 variables
+#' @format A data.frame with 234 observations
 
 "nematode"

### 2022-08-08 R/nvpassoc.R

@@ -35,6 +35,6 @@
 #' @keywords datasets
 #' @name nvpassoc
 #' @usage data(nvpassoc)
-#' @format A data.frame with 7164 observations of 16 variables
+#' @format A data.frame with 7164 observations
 
 "nvpassoc"

### 2022-08-08 R/pathogen.R

@@ -19,6 +19,6 @@
 #' @keywords datasets
 #' @name pathogen
 #' @usage data(pathogen)
-#' @format A data.frame with 2041 observations of 5 variables
+#' @format A data.frame with 2041 observations 
 
 "pathogen"

### 2022-08-08 R/viraref.R

@@ -28,6 +28,6 @@
 #' @keywords datasets
 #' @name viraref
 #' @usage data(viraref)
-#' @format A data.frame with 2124 observations of 5 variables
+#' @format A data.frame with 2124 observations 
 
 "viraref"

### 2022-08-08 R/viruses.R

@@ -7,7 +7,6 @@
 #'   \item ERNv. EDWIP record number
 #'   \item HostSpecies. Host species
 #'   \item VirusType. DNA or RNA virus
-#'   \item Group. Viruses.
 #'   \item PathogenSpecies. Viral family 
 #'   \item Virus. Virus identity
 #'   \item HostStageInfected. Host stage infected
@@ -19,6 +18,7 @@
 #'   \item ModificationDate. Modification date of entry (wrong)
 #'   \item ProvinceA. Canadian provinces where host-virus interaction occurs
 #'   \item PathogenValue. Is there value to the pathogen (can it be used as a control agent?)
+#'   \item Group. Viruses.
 #'   \item HostTaxID. Host NCBI ID number
 #'   \item HostGenus. Host genus
 #'   \item HostFamily. Host family
@@ -36,6 +36,6 @@
 #' @keywords datasets
 #' @name viruses
 #' @usage data(viruses)
-#' @format A data.frame with 1659 observations of 14 variables
+#' @format A data.frame with 1659 observations 
 
 "viruses"

### 2022-08-08 R/zzz.R

@@ -5,28 +5,28 @@
 #' are correct. 
 #' 
 #' \itemize{
-#'   \item ERNnew
-#'   \item HostSpecies
-#'   \item HostOrder
-#'   \item HostFamily
-#'   \item HostHabitat
-#'   \item HostFood
-#'   \item HostGenYr
-#'   \item PathSpecies
-#'   \item PathGroup
-#'   \item PathHighTaxon
-#'   \item PathLowTaxon
-#'   \item StageInf
-#'   \item TissueInfected
-#'   \item Field
-#'   \item Country
-#'   \item IntermediateHost
-#'   \item Citation
-#'   \item MoreInfo
-#'   \item Who
-#'   \item CreationDate
-#'   \item ModificationDate
-#'   \item StainFCB
+#' \item ERNnew. EDWIP record number
+#' \item    HostSpecies. Host species  
+#' \item    HostOrder. Host order  
+#' \item    HostFamily. Host family 
+#' \item    HostHabitat. Habitat type of host 
+#' \item    HostFood. What does the host eat? 
+#' \item    HostGenYr. Number of generations of hosts per year
+#' \item    PathSpecies. Nematode species 
+#' \item    PathGroup. Pathogen group (fungi, protozoa, nematode, etc.) 
+#' \item    PathHighTaxon. Pathogen taxonomic information (mostly `NA`)  
+#' \item    PathLowTaxon. Pathogen taxonomic information (mostly `NA`) 
+#' \item    StageInf. Host life stage infected 
+#' \item    TissueInfected. Host tissue infected 
+#' \item    Field. Was this a field or lab tested association? 
+#' \item    Country. Country of host-pathogen association 
+#' \item    IntermediateHost .  Information on intermediate hosts 
+#' \item    Citation. Citation for host-pathogen record 
+#' \item    MoreInfo. Additional comments or notes  
+#' \item    Who. Identity of researcher who entered data 
+#' \item    CreationDate. Record creation date 
+#' \item    ModificationDate. Record modification date 
+#' \item    StainFCB . takes values: Adult, egg, larvae, pupa   
 #'  }
 #'
 #' @docType data
@@ -45,21 +45,22 @@
 #' 
 #'
 #' \itemize{
-#'   \item ERNnem. EDWIP record number
-#'   \item Host
-#'   \item Nematode
-#'   \item NemaOrder
-#'   \item NemaFamily
-#'   \item NemaStrain
-#'   \item StageInfected
[…]

### 2022-08-23 R/getNCBI.R
par tweaks in vignette and example added for getNCBI

@@ -7,12 +7,14 @@
 #'
 #' @return a data.frame with nrow == length(species) 
 #' @name getNCBI
+#' @examples
+#'   mouse <- getNCBI("Peromyscus leucopus")
+#'   worm  <- getNCBI("Ascaris lumbricoides", host=FALSE)
 #' @importFrom taxize classification get_uid
 #' @importFrom stats na.omit 
 #' @importFrom utils tail
 #' @importFrom plyr mutate
 #' @export
-
 
 
 getNCBI <- function(species, host=TRUE){ 

### 2022-08-25 manuscript/manuscript.tex
fixed the Rbuildignore and the gitignore
updates to manuscript before send off, CRAN badge added to README

@@
 
-%%% Target journal
-% Ecography (software note) -- maybe we shouldn't re-submit and pay $3000 to publish? 
-% Journal of Open Source Software 
-% Ecology (data paper)
-%
-
-%%% Potential reviewers
-% Lewis Bartlett  lewis.Bartlett@uga.edu UGA
-% Philipp Boersch  pboesu@gmail.com   British Trust for Ornithology
-% Ben Longdon b.longdon2@exeter.ac.uk Uni of Exeter
-% Susan Braxton braxton@illinois.edu U Illinois (could be COI with Onstad though)
-% 
-%
-
-%%% Editors
-% Thiago Rangel
-% 
-%
-
 \begin{document}
@@
 \clearpage
-\setstretch{1.35}
+\setstretch{1.25}
 \linenumbers
+
+\noindent \texttt{insectDisease}: programmatic access to the \textit{Ecological Database of the World's Insect Pathogens}
 
@@
 \clearpage
+
 \setstretch{1.9}
@@
 In doing so, we implicitly adhere to the FAIR (Findable, Accessible, Interoperable, Resuable) guidelines for managing data \citep{wilkinson2016}.
-By hosting the data openly on GitHub, and versioning releases of the data with a permanent identifier (DOI), we ensure the longevity and versioned curation of this data resource.
+By hosting the data openly on GitHub (\texttt{https://github.com/viralemergence/insectDisease/}) -- versioning releases of the data with a permanent identifier (DOI) on Zenodo -- and on CRAN (\texttt{https://cran.r-project.org/web/packages/insectDisease/index.html}), we ensure the longevity and versioned curation of this data resource.
 Finally, the incorporation of taxonomic data through \texttt{taxize} \citep{chamberlain2013} ensures that host and pathogen taxonomic names are updated periodically to accommodate for dynamic data or changing taxonomies.
@@
 However, for users who do not wish to download the entire data resource, and simply want to quickly query a static version of the database, there is also a standalone web user interface (\texttt{https://edwip.ecology.uga.edu/}) that allows users to easily subset and explore the data.
-The web interface serves arguably the most important subset of the overall data (data files \texttt{nematode}, \texttt{viruses},\texttt{nvpassoc}, \texttt{negative}, and \texttt{hosts}).
+The web interface serves arguably the most important subset of the overall data (data files \texttt{nematode}, \ \texttt{viruses}, \ \texttt{nvpassoc}, \ \texttt{negative}, and \ \texttt{hosts}).
 This interface allows users to quickly query based on host or parasite taxonomy as a dropdown list.
@@
 GitHub releases of the data ensure versioning is maintained and all versions are accessible.
-At the time of this writing, the current version is 1.2.0 (available at \\ \texttt{https://github.com/viralemergence/insectDisease/releases/tag/1.2.0}).
-Releases are given a DOI through integration with Zenodo (\texttt{https://doi.org/10.5281/zenodo.5821896}).
+At the time of this writing, the current version is 1.2.1 (available at \\ \texttt{https://github.com/viralemergence/insectDisease/releases/tag/1.2.1}).
+\\
+Releases are given a DOI through integration with Zenodo (available at \\ \texttt{https://doi.org/10.5281/zenodo.5821896}).
+The package (release version 1.2.1) is also available on CRAN \\ \texttt{https://cran.r-project.org/web/packages/insectDisease/index.html}.
 
@@
    \texttt{negative} & 529  & 21  & information on negative host-pathogen associations  \\
-   \texttt{nemaref} & 338  & 5  & references from nematode pathogens  \\
+   \texttt{nemaref} & 338  & 16  & references from nematode pathogens  \\
    \texttt{nematode} & 234  & 24  & host-nematode interaction data \\   
@@
    \texttt{viraref} & 2124  & 16  & references from viral infections  \\   
-   \texttt{viruses} & 1659  & 25  & host-viral interaction data  \\      
+   \texttt{viruses} & 1659  & 26  & host-viral interaction data  \\      
   \hline
